---
title: "00 - Global Setup"
author: "Analysis Pipeline"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Global Setup

This notebook defines global variables, loads libraries, and sets up helper functions for the analysis.

## Load Libraries

```{r libraries}
suppressPackageStartupMessages({
  library(tidyverse)
  library(data.table)
  library(phyloseq)
  library(vegan)
  library(zCompositions)
  library(ggplot2)
})
```

## Define Paths

```{r paths}
DATA_DIR <- "/media/Box/applesurface_management/00_Data_antigravity"
RESULTS_DIR <- file.path(DATA_DIR, "results")

if (!dir.exists(RESULTS_DIR)) dir.create(RESULTS_DIR)
if (!dir.exists(file.path(RESULTS_DIR, "data_objects"))) dir.create(file.path(RESULTS_DIR, "data_objects"))
```

## Sample Metadata

12 metagenomes from fruit surfaces:
- ISR_Org (3), ISR_Conv (3), USA_Conv1 (3), USA_Conv2 (3)

```{r metadata}
sample_names <- c(
  "ISR_Org_1", "ISR_Org_2", "ISR_Org_3",
  "ISR_Conv1_1", "ISR_Conv1_2", "ISR_Conv1_3",
  "USA_Conv1_1", "USA_Conv1_2", "USA_Conv1_3",
  "USA_Conv2_1", "USA_Conv2_2", "USA_Conv2_3"
)

metadata <- data.frame(
  SampleID = sample_names,
  Group4 = c(rep("ISR_Org", 3), rep("ISR_Conv", 3), rep("USA_Conv1", 3), rep("USA_Conv2", 3)),
  Country = c(rep("ISR", 6), rep("USA", 6)),
  Management = c(rep("Org", 3), rep("Conv", 3), rep("Conv", 6))
)
rownames(metadata) <- metadata$SampleID

knitr::kable(metadata, caption = "Sample Metadata")
```

## Color Palette

```{r colors}
group4_colors <- c(
  "ISR_Org" = "#1b9e77",
  "ISR_Conv" = "#d95f02",
  "USA_Conv1" = "#7570b3",
  "USA_Conv2" = "#e7298a"
)
```

## Helper Functions

### CLR Transformation

```{r clr_function}
calc_clr <- function(count_matrix) {
  mat <- as.matrix(count_matrix)

  # Zero replacement (multiplicative)
  if (any(mat == 0)) {
    mat_repl <- zCompositions::cmultRepl(mat,
      label = 0, method = "CZM",
      output = "p-counts", suppress.print = TRUE
    )
  } else {
    mat_repl <- mat
  }

  # CLR: log(x) - mean(log(x)) per row
  mat_log <- log(mat_repl)
  mat_clr <- t(apply(mat_log, 1, function(x) x - mean(x)))

  return(mat_clr)
}
```

### Aitchison Distance

```{r aitchison_function}
calc_aitchison <- function(clr_matrix) {
  dist(clr_matrix, method = "euclidean")
}
```

### Feature Filtering

```{r filter_function}
filter_features <- function(count_matrix) {
  total_counts <- colSums(count_matrix)
  presence <- colSums(count_matrix > 0)
  keep <- (total_counts >= 10) & (presence >= 3)
  return(count_matrix[, keep, drop = FALSE])
}
```
### Taxonomy Correction (GTDB)

```{r taxonomy_fix_function}
correct_gtdb_taxonomy <- function(df) {
  # Define known GTDB Archaeal Phyla to fix mislabeling
  gtdb_archaea_phyla <- c(
    "Aenigmatarchaeota", "Altiarchaeota", "Asgardarchaeota",
    "Crenarchaeota", "Halobacteriota", "Methanobacteriota",
    "Thermoplasmatota", "Thermoproteota", "Micrarchaeota",
    "Nanoarchaeota", "Iainarchaeota", "Huberarchaeota",
    "Undinarchaeota", "Nanohaloarchaeota"
  )

  # Ensure columns exist
  if (!all(c("source", "phylum", "kingdom") %in% colnames(df))) {
    warning("Dataframe missing required columns (source, phylum, kingdom). Returning original.")
    return(df)
  }

  df %>%
    dplyr::mutate(
      # 1. Assign Domain (Strict)
      domain = dplyr::case_when(
        # GTDB: Bacteria or Archaea
        source == "GTDB" & (phylum %in% gtdb_archaea_phyla | grepl(paste(gtdb_archaea_phyla, collapse = "|"), lineage)) ~ "Archaea",
        source == "GTDB" ~ "Bacteria",

        # NT Manual Fixes/Inferences
        source == "NT" & grepl("Filasterea|Choanoflagellata|Ichthyosporea", lineage) ~ "Eukaryota",
        source == "NT" & (kingdom %in% c("Fungi", "Metazoa", "Viridiplantae") | grepl("k__Fungi|k__Metazoa|k__Viridiplantae", lineage)) ~ "Eukaryota",

        # NT: Extract d__ if present
        source == "NT" & grepl("d__", lineage) ~ sub(".*d__([^|]+).*", "\\1", lineage),
        # Fallback for standard domains if d__ missing
        source == "NT" & kingdom %in% c("Bacteria", "Archaea", "Viruses") ~ kingdom,
        TRUE ~ "Unclassified"
      ),

      # 2. Assign Kingdom (Hybrid: Kingdom or Domain-as-Kingdom)
      kingdom = dplyr::case_when(
        # GTDB: Duplicate Domain
        source == "GTDB" ~ domain,

        # NT Manual Fixes
        source == "NT" & grepl("Filasterea|Choanoflagellata|Ichthyosporea", lineage) ~ "Protista",

        # NT: Prefer k__ if present
        source == "NT" & grepl("k__", lineage) ~ sub(".*k__([^|]+).*", "\\1", lineage),

        # NT Fallback: If k__ missing, use d__ (Domain) as Kingdom
        source == "NT" & grepl("d__", lineage) ~ sub(".*d__([^|]+).*", "\\1", lineage),

        # NT Fallback: Use existing non-empty Value (even if Domain-like)
        source == "NT" & kingdom != "" & !is.na(kingdom) ~ kingdom,
        TRUE ~ "Unclassified"
      )
    )
}
```

## Save Objects

```{r save}
saveRDS(metadata, file.path(RESULTS_DIR, "metadata.rds"))
saveRDS(group4_colors, file.path(RESULTS_DIR, "group4_colors.rds"))
message("Setup complete.")
```
