---
title: "02 - Alpha Diversity Analysis"
author: "Analysis Pipeline"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 6)
```

# Q2: Alpha Diversity via Hill Numbers

**Question:** Does within-sample diversity differ among Group4 in prokaryotes, plasmids, and viruses?

## Load Setup

```{r load_setup}
source("/media/Box/applesurface_management/00_Data_antigravity/scripts/00_setup.R")
library(emmeans)
library(multcomp)
library(multcompView)

OUT_DIR <- file.path(RESULTS_DIR, "02_alpha_diversity")
if (!dir.exists(OUT_DIR)) dir.create(OUT_DIR)
```

## Hill Number Function

Hill numbers:
- **q=0**: Richness (number of taxa)
- **q=1**: exp(Shannon) - effective number of common taxa
- **q=2**: 1/Simpson - effective number of dominant taxa

```{r hill_function}
calc_hill <- function(count_matrix) {
  rel_abund <- prop.table(as.matrix(count_matrix), 1)

  q0 <- rowSums(rel_abund > 0)
  q1 <- exp(vegan::diversity(rel_abund, index = "shannon"))
  q2 <- 1 / vegan::diversity(rel_abund, index = "simpson")

  data.frame(SampleID = rownames(count_matrix), q0 = q0, q1 = q1, q2 = q2)
}
```

---

# 0. Reads (Rarefied to Equal Depth)

**Scientific Rationale:** Read-level taxonomy from Kraken2 provides the most comprehensive view of community composition. We rarefy to the minimum library size to ensure fair comparisons of alpha diversity across samples.

## Load and Process Read Taxonomy

```{r load_reads}
# Load all read taxonomy files
reads_dir <- file.path(DATA_DIR, "00_reads/Taxonomy")
read_files <- list.files(reads_dir, pattern = "_combined\\.tsv$", full.names = TRUE)

# Read and combine all files
read_data_list <- lapply(read_files, fread)
read_data <- rbindlist(read_data_list, fill = TRUE)

# Filter to valid Bacteria/Archaea (GTDB + NT)
# Filter to valid Bacteria/Archaea/Fungi (GTDB + NT) and normalize
read_data_clean <- read_data %>%
  correct_gtdb_taxonomy() %>%
  mutate(
    # Normalize prefixes (k__, p__, etc.)
    kingdom = gsub("^[kd]__", "", kingdom),
    phylum = gsub("^p__", "", phylum),
    class = gsub("^c__", "", class),
    order = gsub("^o__", "", order),
    family = gsub("^f__", "", family),
    genus = gsub("^g__", "", genus),
    species = gsub("^s__", "", species)
  ) # %>%
# dplyr::filter(kingdom %in% c("Bacteria", "Archaea")) # Removed filter to include all kingdoms

# Function to plot composition with Phylum-based Coloring
# Simple Stacked Barplot Function
# Generic Plotting Function for Composition
plot_composition_bar <- function(df_wide, metadata, rank_col = "DisplayTaxon", metric_label = "Relative Abundance", n_top = 20) {
  # Convert wide to long
  df_long <- df_wide %>%
    rownames_to_column("SampleID") %>%
    pivot_longer(-SampleID, names_to = "Taxon", values_to = "Value") %>%
    left_join(metadata, by = "SampleID")

  # Identify Top Taxa by Mean Value
  top_taxa <- df_long %>%
    group_by(Taxon) %>%
    summarise(MeanVal = mean(Value, na.rm = TRUE)) %>%
    top_n(n_top, MeanVal) %>%
    pull(Taxon)

  # Group "Other"
  df_plot <- df_long %>%
    mutate(DisplayTaxon = ifelse(Taxon %in% top_taxa, Taxon, "Other")) %>%
    group_by(SampleID, Group4, DisplayTaxon) %>%
    summarise(Value = sum(Value, na.rm = TRUE), .groups = "drop")

  # Order Taxa
  df_plot$DisplayTaxon <- factor(df_plot$DisplayTaxon, levels = c(top_taxa, "Other"))

  # Palette
  n_colors <- length(top_taxa)
  my_palette <- colorRampPalette(RColorBrewer::brewer.pal(12, "Paired"))(n_colors)
  names(my_palette) <- top_taxa
  final_palette <- c(my_palette, "Other" = "grey85")

  # Plot
  ggplot(df_plot, aes(x = SampleID, y = Value, fill = DisplayTaxon)) +
    geom_bar(stat = "identity", width = 0.9) +
    facet_grid(~Group4, scales = "free_x", space = "free") +
    scale_fill_manual(values = final_palette) +
    theme_bw(base_size = 12) +
    labs(
      y = metric_label, x = "", fill = "Taxon"
    ) +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10),
      legend.position = "right"
    )
}

cat("## Taxonomic Composition (Reads)\n\n")

# Phylum Level
# Prepare data for Read Plots (Phylum Level)
cat("Read Data Clean Rows:", nrow(read_data_clean), "\n")

if (nrow(read_data_clean) > 0) {
  prep_reads_phylum <- read_data_clean %>%
    filter(!is.na(phylum) & phylum != "") %>%
    group_by(sample, phylum) %>%
    summarise(Count = sum(count, na.rm = TRUE), .groups = "drop")

  if (nrow(prep_reads_phylum) > 0) {
    # Ensure it's a dataframe and ungrouped
    prep_reads_phylum <- as.data.frame(prep_reads_phylum)

    cat("Cols before pivot:", paste(colnames(prep_reads_phylum), collapse = ", "), "\n")

    prep_reads_phylum <- prep_reads_phylum %>%
      pivot_wider(names_from = "phylum", values_from = "Count", values_fill = 0)

    # Check if sample column exists before converting to rownames
    if ("sample" %in% colnames(prep_reads_phylum)) {
      prep_reads_phylum <- prep_reads_phylum %>% column_to_rownames("sample")
    } else {
      stop("Sample column missing after pivot.")
    }

    # Relative Abundance
    prep_reads_phylum_rel <- prop.table(as.matrix(prep_reads_phylum), 1) %>% as.data.frame()

    cat("### Phylum Level: Relative Abundance\n")
    print(plot_composition_bar(prep_reads_phylum_rel, metadata, metric_label = "Relative Abundance (Reads - Phylum)"))

    cat("### Phylum Level: Read Counts\n")
    print(plot_composition_bar(as.data.frame(prep_reads_phylum), metadata, metric_label = "Read Counts (Phylum)"))
  } else {
    message("No phylum data available for reads after summarization.")
  }
} else {
  message("read_data_clean is empty.")
}

# Prepare data for Alpha Diversity (Genus level)
read_genus <- read_data_clean %>%
  filter(!is.na(genus) & genus != "" & genus != "-") %>%
  dplyr::select(genus, count, sample) %>%
  group_by(sample, genus) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop")

# Convert to wide matrix (samples x genus)
read_genus_wide <- read_genus %>%
  pivot_wider(names_from = genus, values_from = count, values_fill = 0) %>%
  column_to_rownames("sample")

# Align with metadata
common_samples_reads <- intersect(rownames(read_genus_wide), rownames(metadata))
read_genus_mat <- as.matrix(read_genus_wide[common_samples_reads, , drop = FALSE])

cat("Samples:", nrow(read_genus_mat), "\n")
cat("Genera:", ncol(read_genus_mat), "\n")
cat("Library sizes:\n")
print(rowSums(read_genus_mat))
```

## Rarefaction

```{r rarefy_reads}
# Load overview mapping to get total sequencing depth
mapping_path <- file.path(DATA_DIR, "overview_mapping.tsv")
overview_map <- as.data.frame(fread(mapping_path))
overview_map <- overview_map %>%
  dplyr::select("Sample", "Total_Reads") %>%
  dplyr::rename(SampleID = "Sample")

# Filter overview map to samples present in data
common_samples <- intersect(rownames(read_genus_mat), overview_map$SampleID)
overview_map <- overview_map %>% filter(SampleID %in% common_samples)
read_genus_mat <- read_genus_mat[common_samples, , drop = FALSE]

# Determine minimum total depth from mapping file
min_total_depth <- min(overview_map$Total_Reads)
cat("Minimum Total Reads (from overview_mapping.tsv):", min_total_depth, "\n")

# Custom rarefaction function using binomial sampling based on total depth
# This accounts for the fact that 'read_genus_mat' is a subset (filtered genera),
# but we want to simulate equal sampling effort based on the ORIGINAL total library size.
rarefy_custom <- function(counts, sample_id, mapping_df, target_depth) {
  sample_total <- mapping_df$Total_Reads[mapping_df$SampleID == sample_id]
  prob <- target_depth / sample_total

  if (prob >= 1) {
    return(counts)
  }

  # Binomial sampling
  rbinom(length(counts), size = counts, prob = prob)
}

# Apply rarefaction
read_rarefied <- read_genus_mat
set.seed(42)

for (i in 1:nrow(read_rarefied)) {
  samp <- rownames(read_rarefied)[i]
  read_rarefied[i, ] <- rarefy_custom(read_rarefied[i, ], samp, overview_map, min_total_depth)
}

# Remove genera with zero counts after rarefaction
read_rarefied <- read_rarefied[, colSums(read_rarefied) > 0]
cat("Genera after rarefaction:", ncol(read_rarefied), "\n")
min_depth <- min_total_depth # For plot title
```

## Taxonomic Composition (Rarefied Reads)

```{r tax_composition_rarefied, fig.width=12, fig.height=7}
# Create Genus -> Phylum mapping from original data
genus_phylum_map <- read_data_clean %>%
  dplyr::filter(!is.na(genus) & genus != "" & !is.na(phylum) & phylum != "") %>%
  dplyr::select(genus, phylum) %>%
  distinct(genus, .keep_all = TRUE) # Keep one mapping per genus

# Convert rarefied matrix (genus counts) to long format
read_rarefied_long <- as.data.frame(read_rarefied) %>%
  tibble::rownames_to_column("SampleID") %>%
  tidyr::pivot_longer(-SampleID, names_to = "genus", values_to = "Count") %>%
  dplyr::filter(Count > 0)

# Join with mapping and summarize by Phylum
rarefied_phylum <- read_rarefied_long %>%
  left_join(genus_phylum_map, by = "genus") %>%
  mutate(phylum = ifelse(is.na(phylum), "Unknown/Unmapped", phylum)) %>%
  group_by(SampleID, phylum) %>%
  summarise(Count = sum(Count), .groups = "drop")

# Convert back to Wide format (Sample x Phylum) for the plotting function
rarefied_phylum_wide <- rarefied_phylum %>%
  tidyr::pivot_wider(names_from = "phylum", values_from = "Count", values_fill = 0) %>%
  tibble::column_to_rownames("SampleID")

# Convert to Relative Abundance
rarefied_phylum_rel <- prop.table(as.matrix(rarefied_phylum_wide), 1) %>% as.data.frame()

# Plot
cat("### Rarefied Reads Composition (Phylum Level)\n")
print(plot_composition_bar(rarefied_phylum_rel, metadata, metric_label = "Relative Abundance (Rarefied)"))
```

## Alpha Diversity (Rarefied)

```{r alpha_helpers}
library(ggpubr)
library(rstatix)

# Enhanced Alpha Diversity Plotting Function (STHDA Style)
plot_hill_stats <- function(hill_data, domain_name, group_col = "Group4", palette = group4_colors) {
  # Reshape to longer format for faceting
  df_long <- hill_data %>%
    pivot_longer(cols = c(q0, q1, q2), names_to = "q", values_to = "Value") %>%
    mutate(q_label = recode(q,
      "q0" = "q=0 (Richness)",
      "q1" = "q=1 (exp(Shannon))",
      "q2" = "q=2 (1/Simpson)"
    )) %>%
    filter(!is.na(Value) & !is.na(.data[[group_col]])) # Remove NAs

  # Ensure group column is a factor
  df_long[[group_col]] <- as.factor(df_long[[group_col]])

  # Check if sufficient data exists
  if (nrow(df_long) == 0) {
    message("No valid data for ", domain_name)
    return(NULL)
  }

  # Perform Pairwise Comparisons (Wilcoxon)
  # Wrap in tryCatch to handle cases with constant values or insufficient groups
  stat.test <- tryCatch(
    {
      df_long %>%
        group_by(q_label) %>%
        wilcox_test(as.formula(paste("Value ~", group_col))) %>%
        adjust_pvalue(method = "BH") %>%
        add_significance("p.adj") %>%
        add_xy_position(x = group_col)
    },
    error = function(e) {
      message("Statistical testing failed for ", domain_name, ": ", e$message)
      return(NULL)
    }
  )

  # Plot
  p <- ggplot(df_long, aes(x = .data[[group_col]], y = Value, fill = .data[[group_col]])) +
    # Jitter points for data distribution
    geom_jitter(width = 0.2, alpha = 0.6, size = 1.5, color = "grey30") +
    # Boxplot on top (optional: make it transparent or use width)
    geom_boxplot(width = 0.5, alpha = 0.8, outlier.shape = NA) +
    facet_wrap(~q_label, scales = "free_y") +
    scale_fill_manual(values = palette) +
    theme_bw(base_size = 12) +
    labs(
      title = paste(domain_name, "Alpha Diversity"),
      y = "Hill Number", x = ""
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.background = element_rect(fill = "grey95")
    )

  # Only add stats if we successfully calculated them and have rows (matches found)
  if (!is.null(stat.test) && nrow(stat.test) > 0) {
    p <- p + stat_pvalue_manual(
      stat.test,
      label = "p.adj.signif",
      tip.length = 0.01,
      hide.ns = TRUE,
      step.increase = 0.1
    )
  }

  return(p)
}

# Statistical Analysis and Plotting Function (LM/GLM/Kruskal)
analyze_and_plot_lm <- function(hill_data, domain_name, group_col = "Group4", palette = group4_colors, model_type = "lm") {
  library(emmeans)
  library(multcomp)
  library(multcompView)
  library(gridExtra)
  library(rstatix)
  library(dplyr)

  # Reshape to longer format
  df_long <- hill_data %>%
    pivot_longer(cols = c(q0, q1, q2), names_to = "q", values_to = "Value") %>%
    mutate(q_label = recode(q,
      "q0" = "q=0 (Richness)",
      "q1" = "q=1 (exp(Shannon))",
      "q2" = "q=2 (1/Simpson)"
    )) %>%
    filter(!is.na(Value) & !is.na(.data[[group_col]]))

  df_long[[group_col]] <- as.factor(df_long[[group_col]])

  # Validate model type
  model_type <- match.arg(model_type, c("lm", "gamma", "kruskal"))

  # Initialize lists
  main_plots <- list()
  diag_plots <- list()

  # Iterate over q indices
  for (cur_q in unique(df_long$q_label)) {
    sub_data <- df_long %>% filter(q_label == cur_q)

    # Check if we have enough data/groups for stats
    n_groups <- length(unique(sub_data[[group_col]]))
    if (n_groups < 2) {
      message(paste("Skipping stats for", cur_q, ": < 2 groups present."))
      p <- ggplot(sub_data, aes(x = .data[[group_col]], y = Value, color = .data[[group_col]])) +
        geom_jitter(width = 0.2) +
        labs(title = cur_q, subtitle = "Not enough groups for stats") +
        theme_bw()
      main_plots[[cur_q]] <- p
      next
    }

    # --- MODEL FITTING & STATS ---

    if (model_type == "kruskal") {
      # Kruskal-Wallis + Dunn's Test
      model_desc <- "Kruskal-Wallis + Dunn's Test"

      # 1. Kruskal Test (Global)
      kw <- kruskal.test(as.formula(paste("Value ~", group_col)), data = sub_data)

      # 2. Dunn's Test (Pairwise)
      dunn <- tryCatch(
        {
          sub_data %>% dunn_test(as.formula(paste("Value ~", group_col)), p.adjust.method = "fdr")
        },
        error = function(e) {
          message(paste("Dunn test failed for", cur_q, ":", e$message))
          return(NULL)
        }
      )

      if (is.null(dunn) || nrow(dunn) == 0) {
        # Fallback
        cld_res <- data.frame(Group = unique(sub_data[[group_col]]), .group = "", stringsAsFactors = FALSE)
        summ <- sub_data %>%
          dplyr::group_by(.data[[group_col]]) %>%
          summarise(MaxVal = max(Value, na.rm = TRUE), Median = median(Value, na.rm = TRUE), .groups = "drop") %>%
          dplyr::rename(Group = all_of(group_col))
        cld_res <- cld_res %>% left_join(summ, by = "Group")
      } else {
        # 3. Generate CLD from Dunn's p-values
        # Create p-value vector for multcompLetters
        # Format: "Group1-Group2" = p.adj
        p_vec <- setNames(dunn$p.adj, paste(dunn$group1, dunn$group2, sep = "-"))

        # multcompLetters expects names like "A-B".
        # dunn_test output group1/group2 are names.

        letters_list <- tryCatch(
          {
            multcompLetters(p_vec, compare = "<", threshold = 0.05, Letters = letters, reversed = TRUE)
          },
          error = function(e) {
            # Fallback if error (e.g. all same)
            list(Letters = setNames(rep("a", length(unique(sub_data[[group_col]]))), unique(sub_data[[group_col]])))
          }
        )

        # Create result dataframe for plotting letters
        # For Kruskal, we usually place letters above the max value or median+spread.
        # Let's compute summary stats (Max) for placement.

        cld_res <- data.frame(
          Group = names(letters_list$Letters),
          Letter = letters_list$Letters,
          stringsAsFactors = FALSE
        )

        # Get Y-positions (Max + offset)
        # Group summaries
        summ <- sub_data %>%
          dplyr::group_by(.data[[group_col]]) %>%
          summarise(MaxVal = max(Value, na.rm = TRUE), Median = median(Value, na.rm = TRUE), .groups = "drop") %>%
          dplyr::rename(Group = all_of(group_col))

        cld_res <- cld_res %>% left_join(summ, by = "Group")

        # Rename column for consistent plotting code below (Map Letter to .group)
        cld_res <- cld_res %>% dplyr::rename(.group = Letter)
      }
    } else {
      # LM or GLM
      if (model_type == "gamma") {
        min_val <- min(sub_data$Value, na.rm = TRUE)
        if (min_val <= 0) {
          # warning(paste("Gamma GLM requires positive values. Adjusted", cur_q))
          sub_data$Value <- sub_data$Value + 1e-6
        }
        mod_fit <- glm(as.formula(paste("Value ~", group_col)), data = sub_data, family = Gamma(link = "log"))
        model_desc <- "Gamma GLM (log link)"
        emm_type <- "response"
      } else {
        mod_fit <- lm(as.formula(paste("Value ~", group_col)), data = sub_data)
        model_desc <- "Linear Model (Gaussian)"
        emm_type <- "response"
      }

      # EMM & CLD
      emm <- emmeans(mod_fit, as.formula(paste("~", group_col)), type = emm_type)

      # FIX: Force as.data.frame to avoid 'rename' error on summary_emm object
      cld_obj <- cld(emm, alpha = 0.05, Letters = letters, adjust = "fdr", reversed = TRUE)
      cld_res <- as.data.frame(cld_obj)
      # Explicitly strip summary_emm class which confuses dplyr::rename
      class(cld_res) <- "data.frame"

      cld_res <- cld_res %>%
        dplyr::rename(Group = all_of(group_col)) %>%
        mutate(q_label = cur_q)

      cld_res$.group <- trimws(cld_res$.group)
      if ("response" %in% colnames(cld_res)) cld_res <- cld_res %>% dplyr::rename(emmean = response)
    }

    # --- PLOTTING ---

    p <- ggplot()

    if (model_type == "kruskal") {
      # Plot: Boxplot + Jitter + Letters (at Max)
      p <- p +
        geom_boxplot(
          data = sub_data, aes(x = .data[[group_col]], y = Value, fill = .data[[group_col]]),
          alpha = 0.6, outlier.shape = NA
        ) +
        geom_jitter(
          data = sub_data, aes(x = .data[[group_col]], y = Value),
          width = 0.2, alpha = 0.4, size = 1.5
        ) +
        geom_text(
          data = cld_res, aes(x = Group, y = MaxVal, label = .group),
          vjust = -0.8, size = 5, fontface = "bold", color = "black"
        ) +
        scale_fill_manual(values = palette) +
        labs(subtitle = paste("Model:", model_desc, "\nLetters: Dunn's Test (BH-adjusted p < 0.05)"))
    } else {
      # Plot: Mean + ErrorBar + Jitter + Letters
      p <- p +
        geom_jitter(
          data = sub_data, aes(x = .data[[group_col]], y = Value, color = .data[[group_col]]),
          width = 0.2, alpha = 0.4, size = 1.5
        ) +
        geom_errorbar(
          data = cld_res, aes(x = Group, ymin = lower.CL, ymax = upper.CL),
          width = 0.2, size = 1, color = "black"
        ) +
        geom_point(
          data = cld_res, aes(x = Group, y = emmean),
          size = 3, color = "black"
        ) +
        geom_text(
          data = cld_res, aes(x = Group, y = upper.CL, label = .group),
          vjust = -0.5, size = 5, fontface = "bold"
        ) +
        scale_color_manual(values = palette) +
        labs(subtitle = paste("Model:", model_desc, "\nError Bars: 95% CI (EMM)\nP-adjust: FDR"))
    }

    # Common Styling
    p <- p +
      facet_wrap(~q_label, scales = "free_y") +
      theme_bw(base_size = 12) +
      labs(y = "Hill Number", x = "", title = cur_q) +
      theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

    main_plots[[cur_q]] <- p

    # --- DIAGNOSTICS ---
    if (model_type != "kruskal") {
      if (model_type == "gamma") {
        resids <- residuals(mod_fit, type = "deviance")
        fitted_vals <- fitted(mod_fit)
        std_resids <- rstandard(mod_fit)
      } else {
        resids <- residuals(mod_fit)
        fitted_vals <- fitted(mod_fit)
        std_resids <- rstandard(mod_fit)
      }

      diag_df <- data.frame(Fitted = fitted_vals, Residuals = resids, StdResid = std_resids)

      p_resid <- ggplot(diag_df, aes(x = Fitted, y = Residuals)) +
        geom_point(alpha = 0.5, color = "blue") +
        geom_hline(yintercept = 0, linetype = "dashed") +
        labs(title = paste("Resid vs Fitted:", cur_q)) +
        theme_gray(base_size = 9)

      p_qq <- ggplot(diag_df, aes(sample = StdResid)) +
        stat_qq(alpha = 0.5) +
        stat_qq_line(color = "red") +
        labs(title = paste("Q-Q Plot:", cur_q)) +
        theme_gray(base_size = 9)

      diag_plots[[cur_q]] <- arrangeGrob(p_resid, p_qq, ncol = 2)
    }
  }

  # Combine
  final_main <- ggarrange(plotlist = main_plots, ncol = 3, nrow = 1, common.legend = TRUE)
  annotated_main <- annotate_figure(final_main, top = text_grob(paste("Analysis:", domain_name), face = "bold", size = 14))

  if (model_type != "kruskal") {
    final_diag <- ggarrange(plotlist = diag_plots, ncol = 1, nrow = 3)
    annotated_diag <- annotate_figure(final_diag, top = text_grob(paste("Diagnostics:", domain_name, "(", model_type, ")"), face = "bold", size = 12))
    return(list(main = annotated_main, diagnostics = annotated_diag))
  } else {
    return(list(main = annotated_main, diagnostics = NULL))
  }
}
```

```{r alpha_reads}
hill_reads <- calc_hill(read_rarefied)
hill_reads$Domain <- "Reads"
hill_reads_meta <- left_join(hill_reads, metadata, by = "SampleID")

# Generate Plot
p_reads <- plot_hill_stats(hill_reads_meta, paste0("Reads (Rarefied to ", format(min_depth, big.mark = ","), ")"))
print(p_reads)
ggsave(file.path(OUT_DIR, "alpha_reads_boxplot.pdf"), p_reads, width = 10, height = 6)
```

## Kruskal-Wallis Tests (Reads)

```{r stats_reads}
stats_reads <- hill_reads_meta %>%
  pivot_longer(cols = c(q0, q1, q2), names_to = "q", values_to = "Value") %>%
  group_by(q) %>%
  summarise(p_value = kruskal.test(Value ~ Group4)$p.value, .groups = "drop")

knitr::kable(stats_reads, caption = "Kruskal-Wallis: Reads Hill Numbers ~ Group4", digits = 4)
```

## Linear Model Analysis (Reads)

```{r lm_reads, fig.width=12, fig.height=9}
# Linear Model (Gaussian)
res_lm_reads <- analyze_and_plot_lm(hill_reads_meta, "Reads", model_type = "lm")
print(res_lm_reads$main)
ggsave(file.path(OUT_DIR, "alpha_reads_lm_main.pdf"), res_lm_reads$main, width = 12, height = 6)

# Diagnostics (Gaussian)
print(res_lm_reads$diagnostics)
ggsave(file.path(OUT_DIR, "alpha_reads_lm_diagnostics.pdf"), res_lm_reads$diagnostics, width = 12, height = 9)

# Gamma GLM (Robust check)
cat("\n### Gamma GLM Check (Reads)\n")
res_gamma_reads <- analyze_and_plot_lm(hill_reads_meta, "Reads", model_type = "gamma")
print(res_gamma_reads$main)
ggsave(file.path(OUT_DIR, "alpha_reads_gamma_main.pdf"), res_gamma_reads$main, width = 12, height = 6)
ggsave(file.path(DATA_DIR, "figures/supplementary/FigS2_alpha_gamma_reads.pdf"), res_gamma_reads$main, width = 12, height = 6)
ggsave(file.path(OUT_DIR, "FigS2_alpha_gamma_reads.pdf"), res_gamma_reads$main, width = 12, height = 6)

# Kruskal-Wallis + Dunn (Non-parametric)
cat("\n### Kruskal-Wallis Check (Reads)\n")
res_kruskal_reads <- analyze_and_plot_lm(hill_reads_meta, "Reads", model_type = "kruskal")
print(res_kruskal_reads$main)
ggsave(file.path(OUT_DIR, "alpha_reads_kruskal_main.pdf"), res_kruskal_reads$main, width = 12, height = 6)
```

---

# 1. Prokaryotes

## Load and Process MAG Data to Species Level

```{r load_mag_data}
# Load CoverM results and taxonomy
mag_counts_raw <- fread(file.path(DATA_DIR, "02_MAGs/coverage/coverm_results.tsv"))
mag_stats <- fread(file.path(DATA_DIR, "02_MAGs/taxonomy/mag_stats_summary_with_taxonomy.tsv"))

# Extract Read Count columns
count_cols <- grep("Read Count", colnames(mag_counts_raw), value = TRUE)
mag_count_mat <- mag_counts_raw %>%
  dplyr::select(Genome, all_of(count_cols)) %>%
  column_to_rownames("Genome")
colnames(mag_count_mat) <- gsub("\\.sorted Read Count", "", colnames(mag_count_mat))

# Parse taxonomy - extract species
mag_tax_parsed <- mag_stats %>%
  dplyr::select(MAG_ID, classification) %>%
  distinct(MAG_ID, .keep_all = TRUE) %>%
  separate(classification,
    into = c("domain", "phylum", "class", "order", "family", "genus", "species"),
    sep = ";"
  ) %>%
  mutate(across(everything(), ~ gsub("^[a-z]__", "", .))) %>%
  mutate(species = ifelse(species == "" | is.na(species), "Unclassified", species))

# Aggregate Read Counts to species level
mag_count_long <- mag_count_mat %>%
  rownames_to_column("Genome") %>%
  pivot_longer(-Genome, names_to = "SampleID", values_to = "Count") %>%
  left_join(mag_tax_parsed, by = c("Genome" = "MAG_ID"))

mag_species_counts <- mag_count_long %>%
  group_by(SampleID, species) %>%
  summarise(Count = sum(Count, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = species, values_from = Count, values_fill = 0) %>%
  column_to_rownames("SampleID")

# Align with metadata and filter
common_samples <- intersect(rownames(mag_species_counts), rownames(metadata))
mag_species_counts <- mag_species_counts[common_samples, , drop = FALSE]
mag_species_counts_filt <- filter_features(mag_species_counts)

cat("Samples:", nrow(mag_species_counts_filt), "\n")
cat("Species:", ncol(mag_species_counts_filt), "\n")

# --- RPKM, TPM, RPM, RPKG Calculations ---

# 1. Load Genome Equivalents for RPKG/RPM
ge_file <- file.path(DATA_DIR, "00_reads/Genome_Equivalents_Summary/ge_summary.csv")
if (file.exists(ge_file)) {
  ge_df <- fread(ge_file)
  # Ensure SampleID column match
  # Assuming colnames are SampleID, Genome_Equivalents, Read_Length, Total_Bases
} else {
  ge_df <- NULL
  message("GE Summary not found. RPKG/RPM calculation limited.")
}

# 2. Extract MAG Lengths
mag_lengths <- mag_stats %>%
  select(MAG_ID, Total_bp) %>%
  column_to_rownames("MAG_ID")
# Ensure alignment with count matrix
common_mags <- intersect(rownames(mag_count_mat), rownames(mag_lengths))
mag_lengths <- mag_lengths[common_mags, , drop = FALSE]
mag_count_mat_calc <- mag_count_mat[common_mags, , drop = FALSE]

# 3. Calculate metrics

# A. RPKM (Existing - from CoverM)
rpkm_cols <- grep("RPKM", colnames(mag_counts_raw), value = TRUE)
mag_rpkm_mat <- mag_counts_raw %>%
  dplyr::select(Genome, all_of(rpkm_cols)) %>%
  column_to_rownames("Genome")
colnames(mag_rpkm_mat) <- gsub("\\.sorted RPKM", "", colnames(mag_rpkm_mat))

# B. TPM (From RPKM)
mag_tpm_mat <- sweep(mag_rpkm_mat, 1, rowSums(mag_rpkm_mat), "/") * 1e6

# C. RPM (Reads Per Million Sequenced Reads)
# If GE data available, use Total_Bases/Read_Length as Total Reads?
# Or just use Sum of Counts (Library Size)?
# User usually implies "Reads Mean" -> Relative Abundance * 1e6.
# But let's verify if they want "Reads Per Million Mapped" (RelAbund*1e6) or "Reads Per Million Sequenced".
# Let's assume "Sequenced" if GE available, else "Mapped".

mag_rpm_mat <- mag_count_mat_calc # Initialize
if (!is.null(ge_df)) {
  # Calculate Total Sequenced Reads per Sample
  # Method robust to missing columns
  cols_needed <- c("SampleID", "Total_Bases", "Read_Length")
  if (all(cols_needed %in% colnames(ge_df))) {
    ge_df$Total_Reads <- ge_df$Total_Bases / ge_df$Read_Length

    # Align samples
    common_samps_ge <- intersect(colnames(mag_rpm_mat), ge_df$SampleID)

    for (s in common_samps_ge) {
      tot_reads <- ge_df[SampleID == s, Total_Reads]
      if (length(tot_reads) > 0 && tot_reads > 0) {
        mag_rpm_mat[, s] <- (mag_count_mat_calc[, s] / tot_reads) * 1e6
      } else {
        mag_rpm_mat[, s] <- NA
      }
    }
    cat("Calculated RPM using Total Sequenced Reads.\n")
  } else {
    # Fallback to Mapped RPM
    mag_rpm_mat <- prop.table(as.matrix(mag_count_mat_calc), 2) * 1e6
    cat("GE Columns missing. Calculated RPM using Mapped Reads (RelAbund * 1e6).\n")
  }
} else {
  mag_rpm_mat <- prop.table(as.matrix(mag_count_mat_calc), 2) * 1e6
  cat("GE file missing. Calculated RPM using Mapped Reads (RelAbund * 1e6).\n")
}


# D. RPKG (Reads Per Kilobase per Genome Equivalent)
# RPKG = (Count / (Len/1000)) / GE
mag_rpkg_mat <- mag_count_mat_calc # Init
if (!is.null(ge_df) && "Genome_Equivalents" %in% colnames(ge_df)) {
  # Calculate RPK first
  mag_len_kb <- mag_lengths$Total_bp / 1000
  mag_rpk_mat <- mag_count_mat_calc / mag_len_kb # Vector recycling? No, mag_len_kb is vector along rows.

  # R expects division by vector to be along COLUMNS by default for matrix?
  # Actually `mat / vec` divides each column by vec? No, it fills column major.
  # use sweep for row-wise division
  mag_rpk_mat <- sweep(mag_count_mat_calc, 1, mag_len_kb, "/")

  # Now divide by GE (Sample-wise)
  common_samps_ge <- intersect(colnames(mag_rpkg_mat), ge_df$SampleID)

  for (s in common_samps_ge) {
    ge_val <- ge_df[SampleID == s, Genome_Equivalents]
    if (length(ge_val) > 0 && ge_val > 0) {
      mag_rpkg_mat[, s] <- mag_rpk_mat[, s] / ge_val
    } else {
      mag_rpkg_mat[, s] <- NA
    }
  }
  cat("Calculated RPKG.\n")
} else {
  mag_rpkg_mat[, ] <- NA
  cat("Cannot calculate RPKG (GE missing).\n")
}


# --- Aggregation to Species Level ---

# Define Helper for Aggregation
agg_to_species <- function(mat, tax_df) {
  if (all(is.na(mat))) {
    return(NULL)
  }

  # Filter common genomes
  comm <- intersect(rownames(mat), tax_df$MAG_ID)
  mat_sub <- mat[comm, , drop = FALSE]
  tax_sub <- tax_df %>% filter(MAG_ID %in% comm)

  # Convert to Long
  mat_long <- mat_sub %>%
    as.data.frame() %>%
    rownames_to_column("MAG_ID") %>%
    pivot_longer(-MAG_ID, names_to = "SampleID", values_to = "Val") %>%
    left_join(tax_sub, by = "MAG_ID")

  # Aggregate
  spec_mat <- mat_long %>%
    group_by(SampleID, species) %>%
    summarise(Val = sum(Val, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = species, values_from = Val, values_fill = 0) %>%
    column_to_rownames("SampleID")

  return(spec_mat)
}

# Aggregate all metrics
mag_species_rpkm <- agg_to_species(mag_rpkm_mat, mag_tax_parsed)
mag_species_tpm <- agg_to_species(mag_tpm_mat, mag_tax_parsed) # Note: Sum of TPMs is valid agg
mag_species_rpm <- agg_to_species(mag_rpm_mat, mag_tax_parsed)
mag_species_rpkg <- agg_to_species(mag_rpkg_mat, mag_tax_parsed)

# Ensure dataframes and filtering (align with metadata)
align_and_prep <- function(mat, meta) {
  if (is.null(mat)) {
    return(NULL)
  }
  comm <- intersect(rownames(mat), rownames(meta))
  if (length(comm) == 0) {
    return(NULL)
  }
  return(as.data.frame(mat[comm, , drop = FALSE]))
}

mag_species_rpkm <- align_and_prep(mag_species_rpkm, metadata)
mag_species_tpm <- align_and_prep(mag_species_tpm, metadata)
mag_species_rpm <- align_and_prep(mag_species_rpm, metadata)
mag_species_rpkg <- align_and_prep(mag_species_rpkg, metadata)

# Common samples aligned for Counts/Rel
mag_species_counts <- mag_species_counts[common_samples, , drop = FALSE]
mag_species_rel <- prop.table(as.matrix(mag_species_counts), 1) %>% as.data.frame()


# Visualize MAG Composition (Species Level)
cat("### Species Level: Relative Abundance\n")
print(plot_composition_bar(mag_species_rel, metadata, metric_label = "Relative Abundance"))

cat("### Species Level: Counts\n")
print(plot_composition_bar(as.data.frame(mag_species_counts), metadata, metric_label = "Read Counts"))

cat("### Species Level: RPKM\n")
if (!is.null(mag_species_rpkm)) print(plot_composition_bar(mag_species_rpkm, metadata, metric_label = "RPKM"))

cat("### Species Level: TPM\n")
if (!is.null(mag_species_tpm)) print(plot_composition_bar(mag_species_tpm, metadata, metric_label = "TPM"))

cat("### Species Level: RPM (Reads Per Million Sequenced)\n")
if (!is.null(mag_species_rpm)) print(plot_composition_bar(mag_species_rpm, metadata, metric_label = "RPM"))

cat("### Species Level: RPKG (Reads Per Kb per GE)\n")
if (!is.null(mag_species_rpkg)) print(plot_composition_bar(mag_species_rpkg, metadata, metric_label = "RPKG"))
```

## Alpha Diversity


```{r alpha_prokaryotes}
hill_mag <- calc_hill(mag_species_counts_filt)
hill_mag$Domain <- "Prokaryotes"
hill_mag_meta <- left_join(hill_mag, metadata, by = "SampleID")

# Generate Plot
p_prok <- plot_hill_stats(hill_mag_meta, "Prokaryotes")
print(p_prok)
ggsave(file.path(OUT_DIR, "alpha_prokaryotes_boxplot.pdf"), p_prok, width = 10, height = 6)
```

## Kruskal-Wallis Tests

```{r stats_prokaryotes}
stats_mag <- hill_mag_meta %>%
  pivot_longer(cols = c(q0, q1, q2), names_to = "q", values_to = "Value") %>%
  group_by(q) %>%
  summarise(p_value = kruskal.test(Value ~ Group4)$p.value, .groups = "drop")

knitr::kable(stats_mag, caption = "Kruskal-Wallis: Prokaryotes Hill Numbers ~ Group4", digits = 4)
```

## Linear Model Analysis (Prokaryotes)

```{r lm_prokaryotes, fig.width=12, fig.height=9}
# Linear Model
res_lm_prok <- analyze_and_plot_lm(hill_mag_meta, "Prokaryotes", model_type = "lm")
print(res_lm_prok$main)
ggsave(file.path(OUT_DIR, "alpha_prokaryotes_lm_main.pdf"), res_lm_prok$main, width = 12, height = 6)

# Diagnostics
print(res_lm_prok$diagnostics)
ggsave(file.path(OUT_DIR, "alpha_prokaryotes_lm_diagnostics.pdf"), res_lm_prok$diagnostics, width = 12, height = 9)

# Gamma GLM
cat("\n### Gamma GLM Check (Prokaryotes)\n")
res_gamma_prok <- analyze_and_plot_lm(hill_mag_meta, "Prokaryotes", model_type = "gamma")
print(res_gamma_prok$main)
ggsave(file.path(OUT_DIR, "alpha_prokaryotes_gamma_main.pdf"), res_gamma_prok$main, width = 12, height = 6)
ggsave(file.path(DATA_DIR, "figures/supplementary/FigS4_alpha_gamma_mags.pdf"), res_gamma_prok$main, width = 12, height = 6)
ggsave(file.path(OUT_DIR, "FigS4_alpha_gamma_mags.pdf"), res_gamma_prok$main, width = 12, height = 6)

# Kruskal-Wallis
cat("\n### Kruskal-Wallis Check (Prokaryotes)\n")
res_kruskal_prok <- analyze_and_plot_lm(hill_mag_meta, "Prokaryotes", model_type = "kruskal")
print(res_kruskal_prok$main)
ggsave(file.path(OUT_DIR, "alpha_prokaryotes_kruskal_main.pdf"), res_kruskal_prok$main, width = 12, height = 6)
```

---

# 2. Viruses

## Composition (Viruses - Top vOTUs)

```{r virus_composition, fig.width=12, fig.height=6}
# Load RPKM data
virus_cov_path <- file.path(DATA_DIR, "04_Viruses/coverage/coverm_results_viruses.tsv")
if (file.exists(virus_cov_path)) {
  virus_cov <- fread(virus_cov_path)

  # Extract RPKM
  rpkm_cols_v <- grep("RPKM", colnames(virus_cov), value = TRUE)
  virus_rpkm_mat <- virus_cov %>%
    dplyr::select(Contig, all_of(rpkm_cols_v)) %>%
    column_to_rownames("Contig")
  colnames(virus_rpkm_mat) <- gsub("\\.sorted RPKM", "", colnames(virus_rpkm_mat))

  # Calculate TPM
  virus_tpm_mat <- sweep(virus_rpkm_mat, 1, rowSums(virus_rpkm_mat), "/") * 1e6

  # Load Counts (already loaded in alpha section usually, but load here for plot)
  virus_counts_path <- file.path(RESULTS_DIR, "data_objects", "virus_counts_filt.rds")
  if (file.exists(virus_counts_path)) {
    virus_counts_rds <- readRDS(virus_counts_path)

    # Align samples
    common_samples_v <- intersect(colnames(virus_rpkm_mat), colnames(virus_counts_rds))
    comm_samp_meta <- intersect(common_samples_v, rownames(metadata))

    if (length(comm_samp_meta) > 0) {
      virus_rpkm_plot <- virus_rpkm_mat[, comm_samp_meta, drop = FALSE]
      virus_tpm_plot <- virus_tpm_mat[, comm_samp_meta, drop = FALSE]
      virus_counts_plot <- virus_counts_rds[, comm_samp_meta, drop = FALSE]

      # Rename rows to vOTU_XXXX if needed?
      # Assuming vOTU IDs are rows.

      # Relative Abundance (from Counts)
      virus_rel <- prop.table(as.matrix(virus_counts_plot), 2) %>% as.data.frame() # NOTE: prop.table margin=2 for columns=samples if input is features x samples
      # Wait, my plot_composition_bar expects features as COLUMNS (wide format) or modifies inside?
      # My plot_composition_bar expects: df_wide (rows=Samples, cols=Taxa)
      # My matrices here are usually Features x Samples (from standard workflow).
      # Let's check mag_species_counts.
      # In `02_alpha_diversity.Rmd`:
      # mag_species_counts was `column_to_rownames("SampleID")` -> So it is Samples x Species.
      # CoverM output `virus_rpkm_mat` is Features x Samples.
      # So I need to TRANSPOSE.

      cat("### Top vOTUs: Relative Abundance\n")
      print(plot_composition_bar(as.data.frame(t(virus_rel)), metadata, metric_label = "Relative Abundance", rank_col = "Taxon", n_top = 20))

      cat("### Top vOTUs: Counts\n")
      print(plot_composition_bar(as.data.frame(t(virus_counts_plot)), metadata, metric_label = "Counts", rank_col = "Taxon", n_top = 20))

      cat("### Top vOTUs: RPKM\n")
      print(plot_composition_bar(as.data.frame(t(virus_rpkm_plot)), metadata, metric_label = "RPKM", rank_col = "Taxon", n_top = 20))

      cat("### Top vOTUs: TPM\n")
      print(plot_composition_bar(as.data.frame(t(virus_tpm_plot)), metadata, metric_label = "TPM", rank_col = "Taxon", n_top = 20))
    }
  }
} else {
  message("Virus coverage file not found.")
}
```

## Alpha Diversity

```{r alpha_viruses}
virus_counts <- readRDS(file.path(RESULTS_DIR, "data_objects", "virus_counts_filt.rds"))

hill_virus <- calc_hill(virus_counts)
hill_virus$Domain <- "Viruses"
hill_virus_meta <- left_join(hill_virus, metadata, by = "SampleID")

# Generate Plot
p_virus <- plot_hill_stats(hill_virus_meta, "Viruses")
print(p_virus)
ggsave(file.path(OUT_DIR, "alpha_viruses_boxplot.pdf"), p_virus, width = 10, height = 6)
```

## Kruskal-Wallis Tests

```{r stats_viruses}
stats_virus <- hill_virus_meta %>%
  pivot_longer(cols = c(q0, q1, q2), names_to = "q", values_to = "Value") %>%
  group_by(q) %>%
  summarise(p_value = kruskal.test(Value ~ Group4)$p.value, .groups = "drop")

knitr::kable(stats_virus, caption = "Kruskal-Wallis: Viruses Hill Numbers ~ Group4", digits = 4)
```

## Linear Model Analysis (Viruses)

```{r lm_viruses, fig.width=12, fig.height=9}
# Linear Model
res_lm_virus <- analyze_and_plot_lm(hill_virus_meta, "Viruses", model_type = "lm")
print(res_lm_virus$main)
ggsave(file.path(OUT_DIR, "alpha_viruses_lm_main.pdf"), res_lm_virus$main, width = 12, height = 6)

# Diagnostics
print(res_lm_virus$diagnostics)
ggsave(file.path(OUT_DIR, "alpha_viruses_lm_diagnostics.pdf"), res_lm_virus$diagnostics, width = 12, height = 9)

# Gamma GLM
cat("\n### Gamma GLM Check (Viruses)\n")
res_gamma_virus <- analyze_and_plot_lm(hill_virus_meta, "Viruses", model_type = "gamma")
print(res_gamma_virus$main)
ggsave(file.path(OUT_DIR, "alpha_viruses_gamma_main.pdf"), res_gamma_virus$main, width = 12, height = 6)
ggsave(file.path(DATA_DIR, "figures/supplementary/FigS5_alpha_gamma_viruses.pdf"), res_gamma_virus$main, width = 12, height = 6)
ggsave(file.path(OUT_DIR, "FigS5_alpha_gamma_viruses.pdf"), res_gamma_virus$main, width = 12, height = 6)

# Kruskal-Wallis
cat("\n### Kruskal-Wallis Check (Viruses)\n")
res_kruskal_virus <- analyze_and_plot_lm(hill_virus_meta, "Viruses", model_type = "kruskal")
print(res_kruskal_virus$main)
ggsave(file.path(OUT_DIR, "alpha_viruses_kruskal_main.pdf"), res_kruskal_virus$main, width = 12, height = 6)
```

---

# 3. Contigs (Genus Level)

```{r alpha_contigs}
contig_counts_path <- file.path(RESULTS_DIR, "data_objects", "contig_genus_counts_filt.rds")

if (file.exists(contig_counts_path)) {
  contig_counts <- readRDS(contig_counts_path)

  hill_contig <- calc_hill(contig_counts)
  hill_contig$Domain <- "Contigs"
  hill_contig_meta <- left_join(hill_contig, metadata, by = "SampleID")

  # Generate Plot
  p_contig <- plot_hill_stats(hill_contig_meta, "Contigs (Genus Level)")
  print(p_contig)
  ggsave(file.path(OUT_DIR, "alpha_contigs_boxplot.pdf"), p_contig, width = 10, height = 6)
} else {
  message("Contig genus counts not found. Run 01_community_composition.Rmd first.")
}
```

## Kruskal-Wallis Tests

```{r stats_contigs}
if (exists("hill_contig_meta")) {
  stats_contig <- hill_contig_meta %>%
    pivot_longer(cols = c(q0, q1, q2), names_to = "q", values_to = "Value") %>%
    group_by(q) %>%
    summarise(p_value = kruskal.test(Value ~ Group4)$p.value, .groups = "drop")

  knitr::kable(stats_contig, caption = "Kruskal-Wallis: Contigs Hill Numbers ~ Group4", digits = 4)
}
```

## Linear Model Analysis (Contigs)

```{r lm_contigs, fig.width=12, fig.height=9}
if (exists("hill_contig_meta")) {
  # Linear Model
  res_lm_contig <- analyze_and_plot_lm(hill_contig_meta, "Contigs", model_type = "lm")
  print(res_lm_contig$main)
  ggsave(file.path(OUT_DIR, "alpha_contigs_lm_main.pdf"), res_lm_contig$main, width = 12, height = 6)
  print(res_lm_contig$diagnostics)
  ggsave(file.path(OUT_DIR, "alpha_contigs_lm_diagnostics.pdf"), res_lm_contig$diagnostics, width = 12, height = 9)

  # Gamma GLM
  cat("\n### Gamma GLM Check (Contigs)\n")
  res_gamma_contig <- analyze_and_plot_lm(hill_contig_meta, "Contigs", model_type = "gamma")
  print(res_gamma_contig$main)
  ggsave(file.path(OUT_DIR, "alpha_contigs_gamma_main.pdf"), res_gamma_contig$main, width = 12, height = 6)
  ggsave(file.path(DATA_DIR, "figures/supplementary/FigS3_alpha_gamma_contigs.pdf"), res_gamma_contig$main, width = 12, height = 6)
  ggsave(file.path(OUT_DIR, "FigS3_alpha_gamma_contigs.pdf"), res_gamma_contig$main, width = 12, height = 6)

  # Kruskal-Wallis
  cat("\n### Kruskal-Wallis Check (Contigs)\n")
  res_kruskal_contig <- analyze_and_plot_lm(hill_contig_meta, "Contigs", model_type = "kruskal")
  print(res_kruskal_contig$main)
  ggsave(file.path(OUT_DIR, "alpha_contigs_kruskal_main.pdf"), res_kruskal_contig$main, width = 12, height = 6)
} else {
  message("Contigs data not available for LM analysis.")
}
```

---

# 4. Plasmids

## Composition (Plasmids - Top Contigs)

```{r plasmid_composition, fig.width=12, fig.height=6}
# Load RPKM data
plasmid_cov_path <- file.path(DATA_DIR, "03_Plasmids/coverage/coverm_results_plasmids.tsv")
if (file.exists(plasmid_cov_path)) {
  plasmid_cov <- fread(plasmid_cov_path)

  # Extract RPKM
  rpkm_cols_p <- grep("RPKM", colnames(plasmid_cov), value = TRUE)
  plasmid_rpkm_mat <- plasmid_cov %>%
    dplyr::select(Contig, all_of(rpkm_cols_p)) %>%
    column_to_rownames("Contig")
  colnames(plasmid_rpkm_mat) <- gsub("\\.sorted RPKM", "", colnames(plasmid_rpkm_mat))

  # Calculate TPM
  plasmid_tpm_mat <- sweep(plasmid_rpkm_mat, 1, rowSums(plasmid_rpkm_mat), "/") * 1e6

  # Extract Counts
  count_cols_p <- grep("Read Count", colnames(plasmid_cov), value = TRUE)
  plasmid_count_mat <- plasmid_cov %>%
    dplyr::select(Contig, all_of(count_cols_p)) %>%
    column_to_rownames("Contig")
  colnames(plasmid_count_mat) <- gsub("\\.sorted Read Count", "", colnames(plasmid_count_mat))

  # Align samples
  common_samples_p <- intersect(colnames(plasmid_rpkm_mat), colnames(plasmid_count_mat))
  comm_samp_meta_p <- intersect(common_samples_p, rownames(metadata))

  if (length(comm_samp_meta_p) > 0) {
    plasmid_rpkm_plot <- plasmid_rpkm_mat[, comm_samp_meta_p, drop = FALSE]
    plasmid_tpm_plot <- plasmid_tpm_mat[, comm_samp_meta_p, drop = FALSE]
    plasmid_counts_plot <- plasmid_count_mat[, comm_samp_meta_p, drop = FALSE]

    # Relative Abundance (from Counts)
    plasmid_rel <- prop.table(as.matrix(plasmid_counts_plot), 2) %>% as.data.frame()

    cat("### Top Plasmid Contigs: Relative Abundance\n")
    print(plot_composition_bar(as.data.frame(t(plasmid_rel)), metadata, metric_label = "Relative Abundance", rank_col = "Taxon", n_top = 20))

    cat("### Top Plasmid Contigs: Counts\n")
    print(plot_composition_bar(as.data.frame(t(plasmid_counts_plot)), metadata, metric_label = "Counts", rank_col = "Taxon", n_top = 20))

    cat("### Top Plasmid Contigs: RPKM\n")
    print(plot_composition_bar(as.data.frame(t(plasmid_rpkm_plot)), metadata, metric_label = "RPKM", rank_col = "Taxon", n_top = 20))

    cat("### Top Plasmid Contigs: TPM\n")
    print(plot_composition_bar(as.data.frame(t(plasmid_tpm_plot)), metadata, metric_label = "TPM", rank_col = "Taxon", n_top = 20))
  }
} else {
  message("Plasmid coverage file not found.")
}
```

## Alpha Diversity (Host Family Level)


```{r alpha_plasmids}
plasmid_counts_path <- file.path(RESULTS_DIR, "data_objects", "plasmid_family_counts_filt.rds")

if (file.exists(plasmid_counts_path)) {
  plasmid_counts <- readRDS(plasmid_counts_path)

  hill_plasmid <- calc_hill(plasmid_counts)
  hill_plasmid$Domain <- "Plasmids"
  hill_plasmid_meta <- left_join(hill_plasmid, metadata, by = "SampleID")

  # Generate Plot
  p_plasmid <- plot_hill_stats(hill_plasmid_meta, "Plasmids (Host Family Level)")
  print(p_plasmid)
  ggsave(file.path(OUT_DIR, "alpha_plasmids_boxplot.pdf"), p_plasmid, width = 10, height = 6)
} else {
  message("Plasmid family counts not found. Run 05_genomic_context.Rmd first.")
}
```

## Kruskal-Wallis Tests

```{r stats_plasmids}
if (exists("hill_plasmid_meta")) {
  stats_plasmid <- hill_plasmid_meta %>%
    pivot_longer(cols = c(q0, q1, q2), names_to = "q", values_to = "Value") %>%
    group_by(q) %>%
    summarise(p_value = kruskal.test(Value ~ Group4)$p.value, .groups = "drop")

  knitr::kable(stats_plasmid, caption = "Kruskal-Wallis: Plasmids Hill Numbers ~ Group4", digits = 4)
}
```

## Linear Model Analysis (Plasmids)

```{r lm_plasmids, fig.width=12, fig.height=5}
if (exists("hill_plasmid_meta")) {
  p_lm_plasmid <- analyze_and_plot_lm(hill_plasmid_meta, "Plasmids")
  print(p_lm_plasmid)
  ggsave(file.path(OUT_DIR, "alpha_plasmids_lm_main.pdf"), p_lm_plasmid$main, width = 12, height = 6)
} else {
  message("Plasmids data not available for LM analysis.")
}
```

---

# Figure 2: Alpha diversity across genomic compartments

This figure visualizes the loss of diversity resolution across genomic compartments, justifying the downstream focus on dominant taxa.

```{r fig2_compartments, fig.width=10, fig.height=8}
# Combine all domains
all_hill <- bind_rows(
  hill_reads_meta %>% dplyr::select(SampleID, Domain, Group4, q0, q1, q2),
  hill_mag_meta %>% dplyr::select(SampleID, Domain, Group4, q0, q1, q2) %>%
    dplyr::mutate(Domain = "MAGs (Species)"),
  hill_virus_meta %>% dplyr::select(SampleID, Domain, Group4, q0, q1, q2)
)

# If these objects exist, also rename their Domain values consistently
if (exists("hill_contig_meta")) {
  all_hill <- bind_rows(
    all_hill,
    hill_contig_meta %>% dplyr::select(SampleID, Domain, Group4, q0, q1, q2)
  )
}

if (exists("hill_plasmid_meta")) {
  all_hill <- bind_rows(
    all_hill,
    hill_plasmid_meta %>% dplyr::select(SampleID, Domain, Group4, q0, q1, q2)
  )
}

# Standardize Domain names (robust even if upstream Domain values differ)
all_hill <- all_hill %>%
  dplyr::mutate(
    Domain = dplyr::case_when(
      Domain %in% c("Reads", "Reads (Genus)", "Read-based", "Read-based (Genus)") ~ "Reads (Genus)",
      Domain %in% c("Contigs", "Contigs (Genus)", "Contig-based", "Contig-based (Genus)") ~ "Contigs (Genus)",
      Domain %in% c("Prokaryotes", "Prokaryotes (Species)", "MAGs", "MAGs (Species)") ~ "MAGs (Species)",
      Domain %in% c("Plasmids", "Plasmids (pOTU)", "pOTU") ~ "Plasmids (pOTU)",
      Domain %in% c("Viruses", "Viruses (vOTU)", "vOTU") ~ "Viruses (vOTU)",
      TRUE ~ as.character(Domain)
    )
  )

# Pivot to long format
all_hill_long <- all_hill %>%
  pivot_longer(cols = c(q0, q1, q2), names_to = "q", values_to = "Value")

# Set factor levels for proper ordering
domain_levels <- c(
  "Reads (Genus)",
  "Contigs (Genus)",
  "MAGs (Species)",
  "Plasmids (pOTU)",
  "Viruses (vOTU)"
)
all_hill_long$Domain <- factor(all_hill_long$Domain, levels = domain_levels)

# Nice facet labels
q_labs <- c(
  "q0" = "Richness (q=0)",
  "q1" = "Shannon (q=1)",
  "q2" = "Simpson (q=2)"
)
all_hill_long$q_label <- factor(
  all_hill_long$q,
  levels = c("q0", "q1", "q2"),
  labels = q_labs
)

# Create Figure 2 - Grid Layout
p_fig2 <- ggplot(all_hill_long, aes(x = Group4, y = Value, fill = Group4)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.4, size = 1) +
  facet_wrap(Domain ~ q_label, scales = "free_y", ncol = 3) +
  theme_bw(base_size = 12) +
  scale_fill_manual(values = group4_colors) +
  labs(
    title = "Alpha Diversity Across Genomic Compartments",
    subtitle = "Each panel has independent Y-axis scale",
    y = "Hill Number",
    x = ""
  ) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey95"),
    strip.text = element_text(size = 9, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7)
  )

print(p_fig2)
ggsave(file.path(DATA_DIR, "figures/main/Fig2_alpha_compartments.pdf"), p_fig2, width = 12, height = 16)
ggsave(file.path(OUT_DIR, "Fig2_alpha_compartments.pdf"), p_fig2, width = 12, height = 16)
```

---

# Figure 3: Concordance Analysis

This figure establishes which diversity signals are preserved across resolutions using correlation analysis.

```{r fig3_concordance, fig.width=10, fig.height=10}
library(ggpubr)

# Helper to prep data pairs
prep_pair <- function(df1, df2, name1, name2) {
  inner_join(
    df1 %>% dplyr::select(SampleID, Group4, q0, q1, q2),
    df2 %>% dplyr::select(SampleID, q0, q1, q2),
    by = "SampleID", suffix = c(paste0("_", name1), paste0("_", name2))
  )
}

plot_list <- list()

# Panel A: Reads vs Contigs (q0)
if (exists("hill_contig_meta")) {
  df_rc <- prep_pair(hill_reads_meta, hill_contig_meta, "Reads", "Contigs")

  p1 <- ggplot(df_rc, aes(x = q0_Reads, y = q0_Contigs)) +
    geom_point(aes(color = Group4), size = 3, alpha = 0.7) +
    geom_smooth(method = "lm", color = "black", se = TRUE, linetype = "dashed") +
    stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top", size = 4) +
    scale_color_manual(values = group4_colors) +
    labs(title = "Reads vs Contigs (q0)", x = "Reads Richness", y = "Contigs Richness") +
    theme_bw(base_size = 12) +
    theme(legend.position = "none")

  plot_list[[length(plot_list) + 1]] <- p1

  # Panel B: Reads vs Contigs (q1)
  p2 <- ggplot(df_rc, aes(x = q1_Reads, y = q1_Contigs)) +
    geom_point(aes(color = Group4), size = 3, alpha = 0.7) +
    geom_smooth(method = "lm", color = "black", se = TRUE, linetype = "dashed") +
    stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top", size = 4) +
    scale_color_manual(values = group4_colors) +
    labs(title = "Reads vs Contigs (q1)", x = "Reads Shannon", y = "Contigs Shannon") +
    theme_bw(base_size = 12) +
    theme(legend.position = "none")

  plot_list[[length(plot_list) + 1]] <- p2
}

# Panel C: Reads vs MAGs (q1)
df_rm <- prep_pair(hill_reads_meta, hill_mag_meta, "Reads", "MAGs")
p3 <- ggplot(df_rm, aes(x = q1_Reads, y = q1_MAGs)) +
  geom_point(aes(color = Group4), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", color = "black", se = TRUE, linetype = "dashed") +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top", size = 4) +
  scale_color_manual(values = group4_colors) +
  labs(title = "Reads vs Prokaryotes (q1)", x = "Reads Shannon", y = "Prokaryotes Shannon") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")

plot_list[[length(plot_list) + 1]] <- p3

# Panel D: Reads vs MAGs (q2)
p4 <- ggplot(df_rm, aes(x = q2_Reads, y = q2_MAGs)) +
  geom_point(aes(color = Group4), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", color = "black", se = TRUE, linetype = "dashed") +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top", size = 4) +
  scale_color_manual(values = group4_colors) +
  labs(title = "Reads vs Prokaryotes (q2)", x = "Reads Simpson", y = "Prokaryotes Simpson") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")

plot_list[[length(plot_list) + 1]] <- p4

# Arrange Grid
p_fig3 <- ggarrange(plotlist = plot_list, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
p_fig3 <- annotate_figure(p_fig3, top = text_grob("Concordance of Alpha Diversity Signals", face = "bold", size = 16))

print(p_fig3)
ggsave(file.path(DATA_DIR, "figures/main/Fig3_alpha_concordance.pdf"), p_fig3, width = 10, height = 10)
```

# 5. Contig Alpha Diversity Normalization Comparison

This section evaluates how different normalization methods affect alpha diversity estimates for Contigs. We compare:
1. **Raw Counts**: Original read counts (sensitive to sequencing depth).
2. **Rarefied Counts**: Subsampled to the minimum library size (standardizes depth, but discards data).
3. **RPKM**: Reads Per Kilobase Million (normalizes for gene length and library size).
4. **Relative Abundance**: Proportion of reads (sum-normalized composition).

```{r contig_norm_comparison, fig.width=10, fig.height=10}
# 1. Load Data for All Methods
# Re-scan for contig files to ensure we get all metric columns
contigs_dir <- file.path(DATA_DIR, "01_contigs/coverage/mapping_metaspades")
file_path_pattern <- "coverm_contigs\\.1000\\.tsv$"
all_files <- list.files(contigs_dir, recursive = TRUE, full.names = TRUE, pattern = file_path_pattern)

if (length(all_files) > 0) {
  read_coverm_metrics <- function(f) {
    dt <- tryCatch(fread(f), error = function(e) NULL)
    if (is.null(dt)) {
      return(NULL)
    }

    # Identify Columns
    id_col <- colnames(dt)[1] # Contig ID
    count_col <- grep("Read Count", colnames(dt), value = TRUE, ignore.case = TRUE)
    rpkm_col <- grep("RPKM", colnames(dt), value = TRUE, ignore.case = TRUE)
    # TPM is the available relative abundance metric
    tpm_col <- grep("TPM", colnames(dt), value = TRUE, ignore.case = TRUE)

    # We need Counts for Rarefaction and RPKM/TPM for comparison
    if (length(count_col) == 0 || length(rpkm_col) == 0 || length(tpm_col) == 0) {
      return(NULL)
    }

    # Extract
    dt_sub <- dt[, c(id_col, count_col, rpkm_col, tpm_col), with = FALSE]
    setnames(dt_sub, c("Contig", "Count", "RPKM", "TPM"))

    # Parse SampleID
    folder_name <- basename(dirname(f))
    clean_sample <- gsub("_contigs\\.1000.*", "", folder_name)
    dt_sub$SampleID <- clean_sample

    return(dt_sub)
  }

  # Load and Combine
  metrics_list <- lapply(all_files, read_coverm_metrics)
  metrics_list <- Filter(Negate(is.null), metrics_list)
  all_metrics <- rbindlist(metrics_list)

  # Filter to Metadata Samples
  all_metrics <- all_metrics[SampleID %in% rownames(metadata)]

  # 2. Reshape into Matrices (Sample x Contig)
  # Raw Counts
  mat_count <- all_metrics %>%
    dplyr::select(SampleID, Contig, Count) %>%
    pivot_wider(names_from = Contig, values_from = Count, values_fill = 0) %>%
    column_to_rownames("SampleID")

  # RPKM
  mat_rpkm <- all_metrics %>%
    dplyr::select(SampleID, Contig, RPKM) %>%
    pivot_wider(names_from = Contig, values_from = RPKM, values_fill = 0) %>%
    column_to_rownames("SampleID")

  # TPM (Relative Abundance)
  mat_tpm <- all_metrics %>%
    dplyr::select(SampleID, Contig, TPM) %>%
    pivot_wider(names_from = Contig, values_from = TPM, values_fill = 0) %>%
    column_to_rownames("SampleID")

  # Rarefied Counts
  min_depth_c <- min(rowSums(mat_count))
  # Only rarefy if depth vary
  if (sd(rowSums(mat_count)) > 0) {
    mat_rarefied <- vegan::rrarefy(mat_count, sample = min_depth_c)
  } else {
    mat_rarefied <- as.matrix(mat_count)
  }

  # 3. Calculate Hill Numbers for Each Method
  calc_hill_norm <- function(mat, method_name) {
    # Ensure matrix format
    mat <- as.matrix(mat)

    res <- tryCatch(
      {
        calc_hill(mat)
      },
      error = function(e) {
        # Fallback if calc_hill fails (e.g. NAs)
        return(NULL)
      }
    )

    if (!is.null(res)) {
      res$Normalization <- method_name
    }
    return(res)
  }

  hill_raw <- calc_hill_norm(mat_count, "Raw Counts")
  hill_rare <- calc_hill_norm(mat_rarefied, paste0("Rarefied (", format(min_depth_c, big.mark = ","), ")"))
  hill_rpkm <- calc_hill_norm(mat_rpkm, "RPKM")
  hill_tpm <- calc_hill_norm(mat_tpm, "TPM (Relative Abundance)")

  # Combine
  hill_compare <- bind_rows(hill_raw, hill_rare, hill_rpkm, hill_tpm)

  if (nrow(hill_compare) > 0) {
    # Add Metadata - Handle SampleID column safely
    meta_for_join <- metadata
    if (!"SampleID" %in% colnames(meta_for_join)) {
      meta_for_join$SampleID <- rownames(meta_for_join)
    }

    hill_compare <- left_join(hill_compare, meta_for_join, by = "SampleID")

    # Reshape for Plotting
    hill_compare_long <- hill_compare %>%
      pivot_longer(cols = c(q0, q1, q2), names_to = "q", values_to = "Value") %>%
      mutate(q_label = factor(q, levels = c("q0", "q1", "q2"), labels = c("Richness (q0)", "Shannon (q1)", "Simpson (q2)")))

    # 4. Plot Comparison
    p_norm_compare <- ggplot(hill_compare_long, aes(x = Normalization, y = Value, fill = Normalization)) +
      geom_boxplot(alpha = 0.7, outlier.shape = NA) +
      geom_jitter(width = 0.2, alpha = 0.4, size = 1) +
      facet_wrap(~q_label, scales = "free_y", ncol = 1) +
      theme_bw(base_size = 12) +
      labs(
        title = "Contig Alpha Diversity: Normalization Comparison",
        subtitle = "Comparing Raw, Rarefied, RPKM, and Relative Abundance",
        y = "Hill Number",
        x = "Normalization Method"
      ) +
      theme(
        legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)
      )


    print(p_norm_compare)
    ggsave(file.path(DATA_DIR, "figures/supplementary/FigS_contig_normalization.pdf"), p_norm_compare, width = 10, height = 10)
    ggsave(file.path(OUT_DIR, "FigS_contig_normalization.pdf"), p_norm_compare, width = 10, height = 10)
  }
} else {
  print("No contig coverage files found for normalization comparison.")
}
```

# 6. Compositional Stability (Stacked Bar Plots)

This section visualizes how normalization affects the taxonomic composition at specific ranks:
- **Reads**: Phylum (Raw vs Rarefied)
- **Contigs**: Genus (Raw, Rarefied, RPKM, TPM)
- **Prokaryotes (MAGs)**: Species (Raw, Rarefied, RPKM)

```{r composition_grid_plot, fig.width=16, fig.height=12}
# Helper Function for Composition Plot
plot_comp_bar <- function(df_long, title, n_top = 15) {
  # df_long must have columns: SampleID, Taxon, Value (Relative Abundance)

  # Identify Top Taxa
  top_taxa <- df_long %>%
    group_by(Taxon) %>%
    summarise(MeanVal = mean(Value, na.rm = TRUE)) %>%
    top_n(n_top, MeanVal) %>%
    pull(Taxon)

  # Group Other
  df_plot <- df_long %>%
    mutate(DisplayTaxon = ifelse(Taxon %in% top_taxa, Taxon, "Other")) %>%
    group_by(SampleID, DisplayTaxon) %>%
    summarise(Value = sum(Value, na.rm = TRUE), .groups = "drop")

  # Order Taxa (Other last)
  df_plot$DisplayTaxon <- factor(df_plot$DisplayTaxon, levels = c(top_taxa, "Other"))

  # Add Metadata
  # Ensure SampleID matches
  meta_for_join <- metadata
  if (!"SampleID" %in% colnames(meta_for_join)) {
    meta_for_join$SampleID <- rownames(meta_for_join)
  }
  df_plot <- left_join(df_plot, meta_for_join, by = "SampleID")

  if (nrow(df_plot) == 0) {
    warning(paste("Plot", title, "has no data after join."))
    return(NULL)
  }

  # Palette
  n_colors <- length(top_taxa)
  # Use a distinct palette
  my_palette <- c(RColorBrewer::brewer.pal(min(n_colors, 12), "Paired"), RColorBrewer::brewer.pal(max(0, n_colors - 12), "Set3"))[1:n_colors]
  names(my_palette) <- top_taxa
  final_palette <- c(my_palette, "Other" = "grey85")

  # Plot
  p <- ggplot(df_plot, aes(x = SampleID, y = Value, fill = DisplayTaxon)) +
    geom_bar(stat = "identity", width = 0.9) +
    facet_grid(~Group4, scales = "free_x", space = "free") +
    scale_fill_manual(values = final_palette) +
    theme_bw(base_size = 10) +
    labs(title = title, y = "Relative Abundance", x = "", fill = "Taxon") +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
      legend.position = "right",
      legend.key.size = unit(0.4, "cm")
    )
  return(p)
}

plot_list_comp <- list()

### 1. Reads (Phylum) ###
if (exists("read_data_clean") && nrow(read_data_clean) > 0) {
  # Raw
  reads_raw_phy <- read_data_clean %>%
    group_by(sample, phylum) %>%
    summarise(Count = sum(count), .groups = "drop") %>%
    group_by(sample) %>%
    mutate(Value = Count / sum(Count)) %>%
    rename(SampleID = sample, Taxon = phylum)

  # Rarefied
  # Map Genus -> Phylum
  genus_phylum_map <- read_data_clean %>%
    select(genus, phylum) %>%
    distinct() %>%
    filter(genus != "" & !is.na(genus))

  if (exists("read_rarefied")) {
    # Rarefied Matrix (Sample x Genus)
    read_rare_long <- as.data.frame(as.matrix(read_rarefied)) %>%
      rownames_to_column("SampleID") %>%
      pivot_longer(-SampleID, names_to = "genus", values_to = "Count") %>%
      left_join(genus_phylum_map, by = "genus") %>%
      filter(!is.na(phylum)) %>%
      group_by(SampleID, phylum) %>%
      summarise(Count = sum(Count), .groups = "drop") %>%
      group_by(SampleID) %>%
      mutate(Value = Count / sum(Count)) %>%
      rename(Taxon = phylum)

    plot_list_comp[["Reads_Rarefied"]] <- plot_comp_bar(read_rare_long, "Reads (Phylum) - Rarefied")
  }

  plot_list_comp[["Reads_Raw"]] <- plot_comp_bar(reads_raw_phy, "Reads (Phylum) - Raw")
}

### 2. Contigs (Phylum) ###
# Load Contig Taxonomy
# Pattern: "01_contigs/Taxonomy/*_combined.tsv"
contig_tax_files <- list.files(file.path(DATA_DIR, "01_contigs/Taxonomy"), pattern = "_combined\\.tsv$", full.names = TRUE)
if (length(contig_tax_files) > 0) {
  contig_tax <- rbindlist(lapply(contig_tax_files, fread), fill = TRUE)
  # Keep ID and Phylum
  if ("phylum" %in% colnames(contig_tax)) {
    contig_tax_map <- contig_tax[, .(contig_id, phylum, sample)]

    if (exists("all_metrics")) {
      # Join with tax
      contig_data <- left_join(all_metrics, contig_tax_map %>% select(contig_id, phylum) %>% distinct(), by = c("Contig" = "contig_id"))

      # Check if we have Phylum
      if (sum(!is.na(contig_data$phylum)) > 0) {
        process_contig_method <- function(col_name, title) {
          contig_data %>%
            filter(!is.na(phylum) & phylum != "") %>%
            group_by(SampleID, phylum) %>%
            summarise(Abundance = sum(.data[[col_name]]), .groups = "drop") %>%
            group_by(SampleID) %>%
            mutate(Value = Abundance / sum(Abundance)) %>%
            rename(Taxon = phylum) %>%
            plot_comp_bar(title)
        }

        plot_list_comp[["Contigs_Raw"]] <- process_contig_method("Count", "Contigs (Phylum) - Raw")
        plot_list_comp[["Contigs_RPKM"]] <- process_contig_method("RPKM", "Contigs (Phylum) - RPKM")
        plot_list_comp[["Contigs_TPM"]] <- process_contig_method("TPM", "Contigs (Phylum) - TPM")

        # Rarefied Contigs
        if (exists("mat_rarefied")) {
          contig_rare_long <- as.data.frame(as.matrix(mat_rarefied)) %>%
            rownames_to_column("SampleID") %>%
            pivot_longer(-SampleID, names_to = "Contig", values_to = "Count") %>%
            left_join(contig_tax_map %>% select(contig_id, phylum) %>% distinct(), by = c("Contig" = "contig_id")) %>%
            filter(!is.na(phylum) & phylum != "") %>%
            group_by(SampleID, phylum) %>%
            summarise(Count = sum(Count), .groups = "drop") %>%
            group_by(SampleID) %>%
            mutate(Value = Count / sum(Count)) %>%
            rename(Taxon = phylum)

          plot_list_comp[["Contigs_Rarefied"]] <- plot_comp_bar(contig_rare_long, "Contigs (Phylum) - Rarefied")
        }
      }
    }
  }
}

### 3. MAGs (Phylum) ###
# Using 'mag_count_mat' (Raw) and 'mag_stats' (Taxonomy)
if (exists("mag_count_mat") && exists("mag_stats")) {
  # Parse MAG Taxonomy (Phylum)
  if (!"Phylum" %in% colnames(mag_stats)) {
    if ("classification" %in% colnames(mag_stats)) {
      mag_stats[, Phylum := gsub(".*p__", "", classification)]
      mag_stats[, Phylum := gsub(";.*", "", Phylum)]
      # Handle cases where p__ is missing (e.g. Unclassified)
      # If no p__, it might retain whole string or be empty. check.
    } else {
      message("MAG Stats missing classification/Phylum column.")
      mag_stats$Phylum <- "Unknown"
    }
  }

  # Map MAG -> Phylum
  # Ensure column names match (MAG_ID vs Genome)
  id_col <- "MAG_ID"
  if (!id_col %in% colnames(mag_stats)) id_col <- "Genome"

  if (id_col %in% colnames(mag_stats)) {
    mag_tax_map <- mag_stats %>%
      select(all_of(id_col), Phylum) %>%
      rename(Genome = all_of(id_col)) %>%
      distinct()

    # Raw
    mag_raw_long <- as.data.frame(as.matrix(mag_count_mat)) %>%
      rownames_to_column("Genome") %>%
      pivot_longer(-Genome, names_to = "SampleID", values_to = "Count") %>%
      left_join(mag_tax_map, by = "Genome") %>%
      filter(!is.na(Phylum)) %>%
      group_by(SampleID, Phylum) %>%
      summarise(Count = sum(Count), .groups = "drop") %>%
      group_by(SampleID) %>%
      mutate(Value = Count / sum(Count)) %>%
      rename(Taxon = Phylum)

    plot_list_comp[["MAGs_Raw"]] <- plot_comp_bar(mag_raw_long, "MAGs (Phylum) - Raw")

    # RPKM logic
    if (exists("mag_counts_raw")) {
      rpkm_cols <- grep("RPKM", colnames(mag_counts_raw), value = TRUE)
      if (length(rpkm_cols) > 0) {
        # SampleID.sorted RPKM
        mag_rpkm_df <- mag_counts_raw %>% select(Genome, all_of(rpkm_cols))

        mag_rpkm_long_raw <- mag_rpkm_df %>%
          pivot_longer(cols = -Genome, names_to = "SampleCol", values_to = "RPKM") %>%
          mutate(SampleID = gsub("\\.sorted RPKM", "", SampleCol)) %>%
          select(-SampleCol)

        mag_rpkm_agg <- mag_rpkm_long_raw %>%
          left_join(mag_tax_map, by = "Genome") %>%
          filter(!is.na(Phylum)) %>%
          group_by(SampleID, Phylum) %>%
          summarise(Val = sum(RPKM), .groups = "drop") %>%
          group_by(SampleID) %>%
          mutate(Value = Val / sum(Val)) %>%
          rename(Taxon = Phylum)

        plot_list_comp[["MAGs_RPKM"]] <- plot_comp_bar(mag_rpkm_agg, "MAGs (Phylum) - RPKM")
      }
    }
  }
}

# Combine Plots
if (length(plot_list_comp) > 0) {
  # Filter out NULLs
  plot_list_comp <- Filter(function(x) !is.null(x), plot_list_comp)

  # Arrange
  p_comp_grid <- ggarrange(plotlist = plot_list_comp, ncol = 2, nrow = ceiling(length(plot_list_comp) / 2), common.legend = FALSE)

  print(p_comp_grid)

  ggsave(file.path(DATA_DIR, "figures/supplementary/FigS_normalization_composition.pdf"), p_comp_grid, width = 20, height = 24, limitsize = FALSE)
  ggsave(file.path(OUT_DIR, "FigS_normalization_composition.pdf"), p_comp_grid, width = 20, height = 24, limitsize = FALSE)

  # Save individual plots as requested
  for (plot_name in names(plot_list_comp)) {
    p <- plot_list_comp[[plot_name]]
    safe_name <- gsub(" ", "_", plot_name)
    # Remove parens etc from filename
    safe_name <- gsub("[()]", "", safe_name)
    fname <- paste0("FigS_comp_", safe_name, ".pdf")
    ggsave(file.path(OUT_DIR, fname), p, width = 10, height = 6)
  }
}

### Summary Composition Plots (Raw, Multiple Taxonomic Levels)


# ===================================================================
# Summary Composition Plots: Loop over Phylum, Class, Order, Family, Genus
# For each level: produce Vertical + Interleaved plots
# ===================================================================

# Load contig taxonomy ONCE (all columns)
if (!exists("contig_tax_full")) {
  contig_tax_files <- list.files(file.path(DATA_DIR, "01_contigs/Taxonomy"), pattern = "_combined\\.tsv$", full.names = TRUE)
  if (length(contig_tax_files) > 0) {
    contig_tax_full <- rbindlist(lapply(contig_tax_files, fread), fill = TRUE)
  }
}

tax_levels <- c("phylum", "class", "order", "family", "genus")

for (tax_level in tax_levels) {
  cat(paste0("\n\n#### ", tools::toTitleCase(tax_level), " Level\n\n"))

  plot_list_summary <- list()
  summary_data_list <- list()

  # --- 1. Reads ---
  if (exists("read_data_clean") && nrow(read_data_clean) > 0 && tax_level %in% colnames(read_data_clean)) {
    reads_tmp <- read_data_clean
    reads_tmp$Taxon <- reads_tmp[[tax_level]]
    reads_agg <- reads_tmp %>%
      filter(!is.na(Taxon) & Taxon != "") %>%
      group_by(sample, Taxon) %>%
      summarise(Count = sum(count), .groups = "drop") %>%
      group_by(sample) %>%
      mutate(Value = Count / sum(Count)) %>%
      rename(SampleID = sample)

    plot_list_summary[["Reads"]] <- plot_comp_bar(reads_agg, "Reads")
    summary_data_list[["Reads"]] <- reads_agg %>% mutate(Domain = "Reads")
  }

  # --- 2. Contigs ---
  if (exists("all_metrics") && exists("contig_tax_full") && tax_level %in% colnames(contig_tax_full)) {
    contig_tax_sub <- contig_tax_full[, c("contig_id", tax_level), with = FALSE]
    colnames(contig_tax_sub)[2] <- "Taxon"
    contig_tax_sub <- unique(contig_tax_sub)
    contig_data_joined <- merge(all_metrics, contig_tax_sub, by.x = "Contig", by.y = "contig_id", all.x = TRUE)

    if (sum(!is.na(contig_data_joined$Taxon)) > 0) {
      contig_agg <- contig_data_joined %>%
        filter(!is.na(Taxon) & Taxon != "") %>%
        group_by(SampleID, Taxon) %>%
        summarise(Abundance = sum(Count), .groups = "drop") %>%
        group_by(SampleID) %>%
        mutate(Value = Abundance / sum(Abundance))

      plot_list_summary[["Contigs"]] <- plot_comp_bar(contig_agg, "Contigs")
      summary_data_list[["Contigs"]] <- contig_agg %>% mutate(Domain = "Contigs")
    }
  }

  # --- 3. MAGs ---
  if (exists("mag_count_mat") && exists("mag_tax_parsed")) {
    if (tax_level %in% colnames(mag_tax_parsed)) {
      mag_tax_sub <- mag_tax_parsed[, c("MAG_ID", tax_level)]
      colnames(mag_tax_sub)[2] <- "Taxon"
      mag_tax_sub <- mag_tax_sub[!is.na(mag_tax_sub$Taxon) & mag_tax_sub$Taxon != "" & mag_tax_sub$Taxon != "Unclassified", ]
      mag_tax_sub <- unique(mag_tax_sub)

      mag_agg <- as.data.frame(as.matrix(mag_count_mat)) %>%
        rownames_to_column("Genome") %>%
        pivot_longer(-Genome, names_to = "SampleID", values_to = "Count") %>%
        merge(mag_tax_sub, by.x = "Genome", by.y = "MAG_ID", all.x = FALSE) %>%
        filter(!is.na(Taxon)) %>%
        group_by(SampleID, Taxon) %>%
        summarise(Count = sum(Count), .groups = "drop") %>%
        group_by(SampleID) %>%
        mutate(Value = Count / sum(Count))

      plot_list_summary[["MAGs"]] <- plot_comp_bar(mag_agg, "MAGs")
      summary_data_list[["MAGs"]] <- mag_agg %>% mutate(Domain = "MAGs")
    }
  }

  # --- Produce Plots ---
  if (length(plot_list_summary) > 0) {
    preferred_order <- c("Reads", "Contigs", "MAGs")
    plot_list_ordered <- plot_list_summary[intersect(preferred_order, names(plot_list_summary))]
    level_title <- tools::toTitleCase(tax_level)

    # A. Vertical Plot
    if (length(plot_list_ordered) > 0) {
      p_vertical <- ggarrange(plotlist = plot_list_ordered, ncol = 1, nrow = length(plot_list_ordered), common.legend = FALSE, align = "v")
      print(p_vertical)
      ggsave(file.path(OUT_DIR, paste0("FigS_summary_composition_vertical_", tax_level, ".pdf")), p_vertical, width = 12, height = 15)
    }

    # B. Interleaved Plot
    if (length(summary_data_list) > 0) {
      df_combined <- bind_rows(summary_data_list)
      df_combined$Domain <- factor(df_combined$Domain, levels = preferred_order)

      meta_sub <- metadata
      if (!"SampleID" %in% colnames(meta_sub)) meta_sub$SampleID <- rownames(meta_sub)
      df_interleaved <- left_join(df_combined, meta_sub, by = "SampleID")

      top_taxa_all <- df_interleaved %>%
        group_by(Taxon) %>%
        summarise(MeanVal = mean(Value, na.rm = TRUE), .groups = "drop") %>%
        slice_max(MeanVal, n = 18) %>%
        pull(Taxon)

      df_interleaved <- df_interleaved %>%
        mutate(DisplayTaxon = ifelse(Taxon %in% top_taxa_all, Taxon, "Other"))
      df_interleaved$DisplayTaxon <- factor(df_interleaved$DisplayTaxon, levels = c(top_taxa_all, "Other"))

      n_cols <- length(top_taxa_all)
      if (n_cols >= 3) {
        pal_base <- RColorBrewer::brewer.pal(min(n_cols, 12), "Paired")
        if (n_cols > 12) pal_base <- c(pal_base, RColorBrewer::brewer.pal(min(n_cols - 12, 12), "Set3"))
        if (n_cols > 24) pal_base <- c(pal_base, RColorBrewer::brewer.pal(min(n_cols - 24, 8), "Dark2"))
        my_pal <- pal_base[1:n_cols]
      } else {
        my_pal <- RColorBrewer::brewer.pal(3, "Paired")[1:n_cols]
      }
      names(my_pal) <- top_taxa_all
      final_pal <- c(my_pal, "Other" = "grey90")

      df_interleaved <- df_interleaved %>%
        arrange(Group4, SampleID, Domain) %>%
        mutate(UniqueBarID = paste(SampleID, Domain, sep = "_"))
      unique_ids <- unique(df_interleaved$UniqueBarID)
      df_interleaved$UniqueBarID <- factor(df_interleaved$UniqueBarID, levels = unique_ids)

      p_interleaved <- ggplot(df_interleaved, aes(x = UniqueBarID, y = Value, fill = DisplayTaxon)) +
        geom_bar(stat = "identity", width = 1) +
        facet_grid(~Group4, scales = "free_x", space = "free") +
        scale_fill_manual(values = final_pal) +
        theme_bw() +
        labs(
          title = paste0("Composition (", level_title, ") - Interleaved"),
          x = "Sample (Reads | Contigs | MAGs)", y = "Relative Abundance"
        ) +
        theme(
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.grid.major.x = element_blank()
        )

      print(p_interleaved)
      ggsave(file.path(OUT_DIR, paste0("FigS_summary_composition_interleaved_", tax_level, ".pdf")), p_interleaved, width = 16, height = 8)
    }
  }
}
```

### Unified-Color Genus Composition (Vertical)

```{r unified_genus_composition, fig.width=16, fig.height=18}
# ===================================================================
# Genus-Level Composition: Same taxa = Same color across compartments
# X-axis labels on bottom only, Domain labels on the right, 45° angle
# ===================================================================

genus_data_list <- list()

# 1. Reads
if (exists("read_data_clean") && nrow(read_data_clean) > 0 && "genus" %in% colnames(read_data_clean)) {
  genus_data_list[["Reads"]] <- read_data_clean %>%
    filter(!is.na(genus) & genus != "") %>%
    group_by(sample, genus) %>%
    summarise(Count = sum(count), .groups = "drop") %>%
    group_by(sample) %>%
    mutate(Value = Count / sum(Count)) %>%
    rename(SampleID = sample, Taxon = genus) %>%
    mutate(Domain = "Reads")
}

# 2. Contigs
if (exists("all_metrics") && exists("contig_tax_full") && "genus" %in% colnames(contig_tax_full)) {
  contig_tax_genus <- contig_tax_full[, .(contig_id, genus)]
  contig_tax_genus <- unique(contig_tax_genus)
  contig_joined <- merge(all_metrics, contig_tax_genus, by.x = "Contig", by.y = "contig_id", all.x = TRUE)
  if (sum(!is.na(contig_joined$genus)) > 0) {
    genus_data_list[["Contigs"]] <- contig_joined %>%
      filter(!is.na(genus) & genus != "") %>%
      group_by(SampleID, genus) %>%
      summarise(Abundance = sum(Count), .groups = "drop") %>%
      group_by(SampleID) %>%
      mutate(Value = Abundance / sum(Abundance)) %>%
      rename(Taxon = genus) %>%
      mutate(Domain = "Contigs")
  }
}

# 3. MAGs
if (exists("mag_count_mat") && exists("mag_tax_parsed") && "genus" %in% colnames(mag_tax_parsed)) {
  mag_tax_genus <- mag_tax_parsed[, c("MAG_ID", "genus")]
  mag_tax_genus <- mag_tax_genus[!is.na(mag_tax_genus$genus) & mag_tax_genus$genus != "" & mag_tax_genus$genus != "Unclassified", ]
  mag_tax_genus <- unique(mag_tax_genus)

  genus_data_list[["MAGs"]] <- as.data.frame(as.matrix(mag_count_mat)) %>%
    rownames_to_column("Genome") %>%
    pivot_longer(-Genome, names_to = "SampleID", values_to = "Count") %>%
    merge(mag_tax_genus, by.x = "Genome", by.y = "MAG_ID", all.x = FALSE) %>%
    filter(!is.na(genus)) %>%
    group_by(SampleID, genus) %>%
    summarise(Count = sum(Count), .groups = "drop") %>%
    group_by(SampleID) %>%
    mutate(Value = Count / sum(Count)) %>%
    rename(Taxon = genus) %>%
    mutate(Domain = "MAGs")
}

if (length(genus_data_list) > 0) {
  # Combine all data
  df_all <- bind_rows(genus_data_list)
  preferred_order <- c("Reads", "Contigs", "MAGs")
  df_all$Domain <- factor(df_all$Domain, levels = preferred_order)
  df_all_raw <- df_all

  # Top genera across ALL domains
  top_genera <- df_all %>%
    group_by(Taxon) %>%
    summarise(MeanVal = mean(Value, na.rm = TRUE), .groups = "drop") %>%
    slice_max(MeanVal, n = 20) %>%
    arrange(desc(MeanVal)) %>%
    pull(Taxon)

  # Classify and group "Other"
  df_all <- df_all %>%
    mutate(DisplayTaxon = ifelse(Taxon %in% top_genera, Taxon, "Other")) %>%
    group_by(SampleID, Domain, DisplayTaxon) %>%
    summarise(Value = sum(Value, na.rm = TRUE), .groups = "drop")
  df_all$DisplayTaxon <- factor(df_all$DisplayTaxon, levels = c(top_genera, "Other"))

  # Build unified palette (safe for n < 3)
  n_top <- length(top_genera)
  if (n_top >= 3) {
    pal_colors <- RColorBrewer::brewer.pal(min(n_top, 12), "Paired")
    if (n_top > 12) pal_colors <- c(pal_colors, RColorBrewer::brewer.pal(min(n_top - 12, 12), "Set3"))
    if (n_top > 24) pal_colors <- c(pal_colors, RColorBrewer::brewer.pal(min(n_top - 24, 8), "Dark2"))
    pal_colors <- pal_colors[1:n_top]
  } else {
    pal_colors <- RColorBrewer::brewer.pal(3, "Paired")[1:n_top]
  }
  names(pal_colors) <- top_genera
  unified_palette <- c(pal_colors, "Other" = "grey85")

  # Add metadata (for Group4 facet)
  meta_sub <- metadata
  if (!"SampleID" %in% colnames(meta_sub)) meta_sub$SampleID <- rownames(meta_sub)
  df_all <- left_join(df_all, meta_sub, by = "SampleID")

  # Single faceted plot: Domain on rows (right labels), Group4 on columns
  p_unified <- ggplot(df_all, aes(x = SampleID, y = Value, fill = DisplayTaxon)) +
    geom_bar(stat = "identity", width = 0.9) +
    facet_grid(Domain ~ Group4, scales = "free_x", space = "free_x", switch = "y") +
    scale_fill_manual(values = unified_palette, drop = FALSE) +
    guides(fill = guide_legend(ncol = 1)) +
    theme_bw(base_size = 10) +
    labs(y = "Relative Abundance", x = "", fill = "Genus") +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 7),
      strip.text.y.right = element_text(size = 11, face = "bold"),
      strip.text.x = element_text(size = 10),
      strip.background = element_rect(fill = "grey95"),
      legend.position = "right",
      legend.key.size = unit(0.6, "cm"),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14, face = "bold"),
      panel.spacing.y = unit(0.3, "cm")
    )

  print(p_unified)

  ggsave(file.path(OUT_DIR, "FigS_summary_composition_vertical_genus_unified.pdf"),
    p_unified,
    width = 16, height = 18
  )

  # Build alternative palette
  if (n_top >= 3) {
    alt_colors <- RColorBrewer::brewer.pal(min(n_top, 12), "Set3")
    if (n_top > 12) alt_colors <- c(alt_colors, RColorBrewer::brewer.pal(min(n_top - 12, 8), "Dark2"))
    if (n_top > 20) alt_colors <- c(alt_colors, RColorBrewer::brewer.pal(min(n_top - 20, 12), "Paired"))
    alt_colors <- alt_colors[1:n_top]
  } else {
    alt_colors <- RColorBrewer::brewer.pal(3, "Set3")[1:n_top]
  }
  names(alt_colors) <- top_genera
  unified_palette_alt <- c(alt_colors, "Other" = "black")

  p_unified_alt <- p_unified + scale_fill_manual(values = unified_palette_alt, drop = FALSE)

  ggsave(file.path(OUT_DIR, "FigS_summary_composition_vertical_genus_unified_alt_colors.pdf"),
    p_unified_alt,
    width = 16, height = 18
  )

  # --- VERSION 3: Top 25 Genera ---
  top_genera_25 <- df_all_raw %>%
    group_by(Taxon) %>%
    summarise(MeanVal = mean(Value, na.rm = TRUE), .groups = "drop") %>%
    slice_max(MeanVal, n = 25) %>%
    arrange(desc(MeanVal)) %>%
    pull(Taxon)

  df_all_25 <- df_all_raw %>%
    mutate(DisplayTaxon = ifelse(Taxon %in% top_genera_25, Taxon, "Other")) %>%
    group_by(SampleID, Domain, DisplayTaxon) %>%
    summarise(Value = sum(Value, na.rm = TRUE), .groups = "drop")
  df_all_25$DisplayTaxon <- factor(df_all_25$DisplayTaxon, levels = c(top_genera_25, "Other"))

  n_top_25 <- length(top_genera_25)
  if (n_top_25 >= 3) {
    pal_colors_25 <- RColorBrewer::brewer.pal(min(n_top_25, 12), "Paired")
    if (n_top_25 > 12) pal_colors_25 <- c(pal_colors_25, RColorBrewer::brewer.pal(min(n_top_25 - 12, 12), "Set3"))
    if (n_top_25 > 24) pal_colors_25 <- c(pal_colors_25, RColorBrewer::brewer.pal(min(n_top_25 - 24, 8), "Dark2"))
    pal_colors_25 <- pal_colors_25[1:n_top_25]
  } else {
    pal_colors_25 <- RColorBrewer::brewer.pal(3, "Paired")[1:n_top_25]
  }
  names(pal_colors_25) <- top_genera_25
  unified_palette_25 <- c(pal_colors_25, "Other" = "grey85")

  df_all_25 <- left_join(df_all_25, meta_sub, by = "SampleID")

  p_unified_25 <- ggplot(df_all_25, aes(x = SampleID, y = Value, fill = DisplayTaxon)) +
    geom_bar(stat = "identity", width = 0.9) +
    facet_grid(Domain ~ Group4, scales = "free_x", space = "free_x", switch = "y") +
    scale_fill_manual(values = unified_palette_25, drop = FALSE) +
    guides(fill = guide_legend(ncol = 1)) +
    theme_bw(base_size = 10) +
    labs(y = "Relative Abundance", x = "", fill = "Genus") +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 7),
      strip.text.y.right = element_text(size = 11, face = "bold"),
      strip.text.x = element_text(size = 10),
      strip.background = element_rect(fill = "grey95"),
      legend.position = "right",
      legend.key.size = unit(0.6, "cm"),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14, face = "bold"),
      panel.spacing.y = unit(0.3, "cm")
    )

  ggsave(file.path(OUT_DIR, "FigS_summary_composition_vertical_genus_unified_25taxa.pdf"),
    p_unified_25,
    width = 16, height = 18
  )
}
```
---

# Session Info

```{r session}
sessionInfo()
```

