---
title: "03 - Differential Abundance Analysis"
author: "Analysis Pipeline"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6
)
options(width = 120)

# Reliable rendering (avoids fragile RStudio notebook cache):
# rmarkdown::render("03_differential_abundance.Rmd", clean = TRUE)
```

# Differential Abundance (Prokaryotes, Plasmids, Viruses)

**Descriptive:** differences among **Group4** (composite groups; confounded).
**Inferential contrasts:**

* **Management within Israel:** Org vs Conv
* **Country within conventional only:** ISR vs USA

---

## Load setup + data

```{r load_setup}
source("/media/Box/applesurface_management/00_Data_antigravity/scripts/00_setup.R")

OUT_DIR <- file.path(RESULTS_DIR, "03_differential_abundance")
dir.create(OUT_DIR, recursive = TRUE, showWarnings = FALSE)

stopifnot(exists("metadata"))
stopifnot(all(c("SampleID", "Group4", "Country", "Management") %in% colnames(metadata)))
if (!all(rownames(metadata) == metadata$SampleID)) rownames(metadata) <- metadata$SampleID
```

```{r load_data}
mag_counts <- readRDS(file.path(RESULTS_DIR, "data_objects", "mag_species_counts_filt.rds"))
cat("Loaded MAG counts:", nrow(mag_counts), "samples x", ncol(mag_counts), "species\n")

plasmid_counts_path <- file.path(RESULTS_DIR, "data_objects", "plasmid_family_counts_filt.rds")
plasmid_counts <- if (file.exists(plasmid_counts_path)) readRDS(plasmid_counts_path) else NULL
if (!is.null(plasmid_counts)) cat("Loaded plasmid counts:", nrow(plasmid_counts), "samples x", ncol(plasmid_counts), "families\n")

virus_counts_path <- file.path(RESULTS_DIR, "data_objects", "virus_counts_filt.rds")
virus_counts <- if (file.exists(virus_counts_path)) readRDS(virus_counts_path) else NULL
if (!is.null(virus_counts)) cat("Loaded virus counts:", nrow(virus_counts), "samples x", ncol(virus_counts), "vOTUs\n")

# --- Rename vOTUs to human-readable short names (vOTU_001, vOTU_002, ...) ---
# Ordered by total abundance (most abundant first)
if (!is.null(virus_counts)) {
  total_ab <- colSums(virus_counts)
  rank_order <- order(total_ab, decreasing = TRUE)

  votu_mapping <- data.frame(
    original_id = colnames(virus_counts)[rank_order],
    votu_name = sprintf("vOTU_%04d", seq_along(rank_order)),
    total_abundance = total_ab[rank_order],
    stringsAsFactors = FALSE
  )

  # Save the mapping table
  write.table(votu_mapping, file.path(OUT_DIR, "votu_name_mapping.tsv"),
    sep = "\t", row.names = FALSE, quote = FALSE
  )
  cat("Saved vOTU name mapping to:", file.path(OUT_DIR, "votu_name_mapping.tsv"), "\n")

  # Apply the renaming to virus_counts
  name_lookup <- setNames(votu_mapping$votu_name, votu_mapping$original_id)
  colnames(virus_counts) <- name_lookup[colnames(virus_counts)]
  cat("Renamed", ncol(virus_counts), "vOTUs (vOTU_0001 = most abundant)\n")
}
```

```{r sanity_checks}
cat("Group4:\n")
print(table(metadata$Group4, useNA = "ifany"))
cat("Country:\n")
print(table(metadata$Country, useNA = "ifany"))
cat("Management:\n")
print(table(metadata$Management, useNA = "ifany"))

stopifnot(length(intersect(rownames(mag_counts), rownames(metadata))) == nrow(metadata))
stopifnot(all(mag_counts >= 0))

if (!is.null(plasmid_counts)) {
  stopifnot(length(intersect(rownames(plasmid_counts), rownames(metadata))) == nrow(metadata))
  stopifnot(all(plasmid_counts >= 0))
}
if (!is.null(virus_counts)) {
  stopifnot(length(intersect(rownames(virus_counts), rownames(metadata))) == nrow(metadata))
  stopifnot(all(virus_counts >= 0))
}
```

---

# 1) Top taxa overview (Prokaryotes)

```{r top_taxa_plot}
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(tibble)
})

rel_ab <- prop.table(as.matrix(mag_counts), 1)
mean_ab <- colMeans(rel_ab)
top_taxa <- names(sort(mean_ab, decreasing = TRUE)[1:min(15, length(mean_ab))])

plot_df <- as.data.frame(rel_ab[, top_taxa, drop = FALSE]) %>%
  tibble::rownames_to_column("SampleID") %>%
  tidyr::pivot_longer(-SampleID, names_to = "Species", values_to = "RelAbund") %>%
  dplyr::left_join(metadata, by = "SampleID")

species_order <- plot_df %>%
  group_by(Species) %>%
  summarise(m = mean(RelAbund), .groups = "drop") %>%
  arrange(desc(m)) %>%
  pull(Species)

p_top <- ggplot(plot_df, aes(x = factor(Species, levels = species_order), y = RelAbund, fill = Group4)) +
  geom_boxplot(position = position_dodge(width = 0.8), outlier.size = 0.6) +
  scale_fill_manual(values = group4_colors) +
  theme_bw(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top 15 Prokaryotic Species by Management \u00d7 Country", x = "Species", y = "Relative abundance")

print(p_top)
ggsave(file.path(OUT_DIR, "top_species_by_group4.pdf"), p_top, width = 10, height = 6)
```

---

# 2) Contrasts

```{r define_contrasts}
suppressPackageStartupMessages(library(rlang))

make_contrast <- function(name, filter_expr, design_col, levels = NULL, note = NULL) {
  list(
    name = name,
    filter_expr = rlang::enquo(filter_expr),
    design_col = design_col,
    levels = levels,
    note = note
  )
}

contrasts <- list(
  make_contrast(
    "Global_Group4", TRUE, "Group4", NULL,
    "Descriptive only (composite groups; confounded)."
  ),
  make_contrast(
    "Israel_Management_Org_vs_Conv", Country == "ISR", "Management", c("Org", "Conv"),
    "Inferential: management effect within Israel."
  ),
  make_contrast(
    "Conventional_Country_ISR_vs_USA", Management == "Conv", "Country", c("ISR", "USA"),
    "Inferential: country effect within conventional only."
  )
)

print(lapply(contrasts, function(x) x[c("name", "design_col", "levels", "note")]))
```

---

# 3) Helpers: filtering, CLR, orchard parsing

```{r helpers}
pick_prev_frac <- function(n) {
  if (n <= 6) {
    return(3 / 6)
  }
  if (n <= 9) {
    return(4 / 9)
  }
  return(0.3)
}

domain_filter_params <- function(domain_name, n_samples) {
  if (grepl("Viruses", domain_name, ignore.case = TRUE)) {
    list(min_prev_frac = if (n_samples <= 9) 4 / 9 else 0.3, min_total = 20)
  } else {
    list(min_prev_frac = pick_prev_frac(n_samples), min_total = 10)
  }
}

filter_counts <- function(counts_df, meta_sub, min_prev_frac = 0.3, min_total = 10) {
  x <- counts_df[rownames(meta_sub), , drop = FALSE]
  x <- x[, colSums(x) > 0, drop = FALSE]
  if (ncol(x) == 0) {
    return(x)
  }

  prev <- colMeans(x > 0)
  keep_prev <- prev >= min_prev_frac
  keep_total <- colSums(x) >= min_total
  x[, keep_prev & keep_total, drop = FALSE]
}

summarize_matrix_qc <- function(counts_df, meta_sub) {
  x <- counts_df[rownames(meta_sub), , drop = FALSE]
  data.frame(
    n_samples = nrow(x),
    n_features = ncol(x),
    libsize_min = min(rowSums(x)),
    libsize_median = median(rowSums(x)),
    libsize_max = max(rowSums(x)),
    zero_frac_median = median(rowMeans(x == 0)),
    zero_frac_max = max(rowMeans(x == 0))
  )
}

# CLR transform (compositions-aware); returns samples x features matrix
clr_transform <- function(counts_df_samples_x_features, pseudocount = 1e-6) {
  suppressPackageStartupMessages(library(compositions))
  x <- as.matrix(counts_df_samples_x_features)
  rs <- rowSums(x)
  if (any(rs == 0)) stop("Some samples have zero total counts after filtering.")
  x <- x / rs
  x[x == 0] <- pseudocount
  compositions::clr(x)
}

# Heuristic orchard ID from SampleID:
# strips terminal replicate suffix (_1/_2/_3), keeps remaining prefix.
infer_orchard <- function(sample_id) {
  sub("_[0-9]+$", "", sample_id)
}

# Human-readable domain names for plot titles
pretty_domain <- function(d) {
  switch(d,
    "Prokaryotes_MAG_species" = "MAGs (Species)",
    "Plasmids_host_family"    = "Plasmid Host Families",
    "Viruses_vOTU"            = "Viral OTUs",
    d # fallback
  )
}

# Human-readable contrast names for plot titles
pretty_contrast <- function(cc_name) {
  switch(cc_name,
    "Global_Group4"                     = "All Groups",
    "Israel_Management_Org_vs_Conv"     = "Organic vs. Conventional (Israel)",
    "Conventional_Country_ISR_vs_USA"   = "Israel vs. USA (Conventional)",
    gsub("_", " ", cc_name) # fallback
  )
}
```

---

# 4) Community plots + PERMANOVA (with dispersion) + optional stratified PERMANOVA

```{r community_tests}
plot_aitchison_pcoa <- function(counts_df, meta_sub, color_col, title, palette_named = NULL) {
  suppressPackageStartupMessages(library(ggplot2))
  clr <- clr_transform(counts_df[rownames(meta_sub), , drop = FALSE])
  pcoa <- cmdscale(dist(clr), k = 2, eig = TRUE)

  df <- data.frame(
    SampleID = rownames(meta_sub),
    PC1 = pcoa$points[, 1],
    PC2 = pcoa$points[, 2],
    Group = meta_sub[[color_col]]
  )

  p <- ggplot(df, aes(PC1, PC2, color = Group)) +
    geom_point(size = 3) +
    stat_ellipse(type = "t", linetype = "dashed", level = 0.68) +
    theme_bw() +
    labs(
      title = title,
      x = paste0("PC1 (", round(100 * pcoa$eig[1] / sum(pcoa$eig), 1), "%)"),
      y = paste0("PC2 (", round(100 * pcoa$eig[2] / sum(pcoa$eig), 1), "%)")
    )
  if (!is.null(palette_named)) p <- p + scale_color_manual(values = palette_named)
  p
}

run_aitchison_permanova <- function(counts_df, meta_sub, design_col, title_tag, output_dir, nperm = 999) {
  suppressPackageStartupMessages(library(vegan))
  clr <- clr_transform(counts_df[rownames(meta_sub), , drop = FALSE])
  dist_mat <- dist(clr)

  if (length(unique(meta_sub[[design_col]])) < 2) {
    return(NULL)
  }

  ad <- vegan::adonis2(dist_mat ~ meta_sub[[design_col]], permutations = nperm)
  bd <- vegan::betadisper(dist_mat, meta_sub[[design_col]])
  bd_test <- vegan::permutest(bd, permutations = nperm)

  out <- list(adonis = ad, betadisper = bd_test, nperm_requested = nperm)
  saveRDS(out, file.path(output_dir, paste0("permanova_betadisper_", title_tag, ".rds")))
  out
}

# Orchard-stratified PERMANOVA (reviewer-friendly).
# Only runs if Orchard has >=2 strata AND each stratum has >=2 samples overall.
run_stratified_permanova <- function(counts_df, meta_sub, design_col, strata_col, title_tag, output_dir, nperm = 999) {
  suppressPackageStartupMessages(library(vegan))

  if (!(strata_col %in% colnames(meta_sub))) {
    return(NULL)
  }
  strata <- meta_sub[[strata_col]]

  # Need at least 2 strata and some within-stratum replication
  if (length(unique(strata)) < 2) {
    return(NULL)
  }
  if (any(table(strata) < 2)) {
    return(NULL)
  }

  clr <- clr_transform(counts_df[rownames(meta_sub), , drop = FALSE])
  dist_mat <- dist(clr)

  # permutations within strata
  ad_str <- vegan::adonis2(dist_mat ~ meta_sub[[design_col]], permutations = nperm, strata = strata)

  out <- list(adonis_stratified = ad_str, strata_col = strata_col, nperm_requested = nperm)
  saveRDS(out, file.path(output_dir, paste0("permanova_stratified_", title_tag, ".rds")))
  out
}
```

---

# 5) Prevalence enrichment (pairwise) + plot

```{r prevalence}
run_prevalence_enrichment <- function(counts_df, meta_sub, group_col, g1, g2, detect_thresh = 1) {
  suppressPackageStartupMessages({
    library(dplyr)
  })

  meta_pair <- meta_sub %>% filter(.data[[group_col]] %in% c(g1, g2))
  if (nrow(meta_pair) < 4) {
    return(NULL)
  }

  x <- counts_df[rownames(meta_pair), , drop = FALSE]
  x <- x[, colSums(x) > 0, drop = FALSE]
  if (ncol(x) == 0) {
    return(NULL)
  }

  y <- meta_pair[[group_col]]
  present <- (x >= detect_thresh) * 1

  res <- lapply(colnames(present), function(feat) {
    tab <- table(y, present[, feat])
    if (!all(c(g1, g2) %in% rownames(tab))) {
      return(NULL)
    }
    if (!all(c(0, 1) %in% colnames(tab))) {
      return(NULL)
    }

    tab <- tab[c(g1, g2), c("0", "1"), drop = FALSE]
    ft <- fisher.test(tab)

    prev_g1 <- mean(present[y == g1, feat] == 1)
    prev_g2 <- mean(present[y == g2, feat] == 1)

    data.frame(
      feature = feat,
      prev_g1 = prev_g1,
      prev_g2 = prev_g2,
      prev_diff = prev_g1 - prev_g2,
      odds_ratio = unname(ft$estimate),
      p = ft$p.value
    )
  }) %>% bind_rows()

  if (is.null(res) || nrow(res) == 0) {
    return(NULL)
  }
  res$q <- p.adjust(res$p, method = "BH")
  res[order(res$q, -abs(res$prev_diff)), ]
}

plot_prevalence_shift <- function(prev_df, title, top_n = 20) {
  suppressPackageStartupMessages(library(ggplot2))
  if (is.null(prev_df) || nrow(prev_df) == 0) {
    return(NULL)
  }

  df <- head(prev_df, top_n)
  ggplot(df, aes(x = prev_diff, y = reorder(feature, prev_diff))) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_point(aes(size = abs(prev_diff), color = prev_diff > 0), alpha = 0.8) +
    theme_bw() +
    labs(title = title, x = "Prevalence Difference", y = "") +
    guides(color = "none")
}
```

---

# 6) Feature-level DA: ANCOM-BC2 + optional ALDEx2 (pairwise only)

```{r da_engines}
pick_ancom_coef <- function(res_df) {
  q_cols <- grep("^q_", colnames(res_df), value = TRUE)
  q_cols_nonint <- setdiff(q_cols, "q_(Intercept)")
  q_pick <- if (length(q_cols_nonint) > 0) q_cols_nonint[1] else q_cols[1]
  sub("^q_", "", q_pick)
}

plot_ancombc_bars <- function(ancom_obj, title_tag, contrast_level = NULL, top_n = 20) {
  suppressPackageStartupMessages({
    library(dplyr)
    library(ggplot2)
  })
  if (is.null(ancom_obj) || is.null(ancom_obj$res)) {
    return(NULL)
  }
  res <- ancom_obj$res
  if (!("taxon" %in% colnames(res))) {
    return(NULL)
  }

  if (is.null(contrast_level)) contrast_level <- pick_ancom_coef(res)

  lfc_col <- paste0("lfc_", contrast_level)
  q_col <- paste0("q_", contrast_level)
  if (!(lfc_col %in% colnames(res)) || !(q_col %in% colnames(res))) {
    return(NULL)
  }

  sig_df <- res %>%
    filter(!is.na(.data[[q_col]]), .data[[q_col]] < 0.05) %>%
    arrange(desc(abs(.data[[lfc_col]]))) %>%
    head(top_n)
  if (nrow(sig_df) == 0) {
    return(NULL)
  }

  ggplot(sig_df, aes(x = reorder(taxon, .data[[lfc_col]]), y = .data[[lfc_col]], fill = .data[[lfc_col]] > 0)) +
    geom_col() +
    coord_flip() +
    scale_fill_manual(values = c("TRUE" = "#2171B5", "FALSE" = "#CB181D"), labels = c("Depleted", "Enriched")) +
    theme_bw() +
    labs(title = paste0(title_tag, " \u2014 Differentially Abundant Taxa"), x = "", y = "Log Fold Change") +
    guides(fill = guide_legend(title = "Direction"))
}

run_aldex2_pairwise <- function(counts_mat_features_x_samples, group_vec, mc.samples = 128, denom = "all") {
  stopifnot(requireNamespace("ALDEx2", quietly = TRUE))
  group_vec <- as.factor(group_vec)
  if (nlevels(group_vec) != 2) {
    return(NULL)
  }

  clr <- ALDEx2::aldex.clr(counts_mat_features_x_samples, group_vec,
    mc.samples = mc.samples, denom = denom, verbose = FALSE
  )
  tt <- ALDEx2::aldex.ttest(clr, verbose = FALSE)
  ef <- ALDEx2::aldex.effect(clr, verbose = FALSE)
  res <- as.data.frame(cbind(tt, ef))
  res$taxon <- rownames(res)
  rownames(res) <- NULL
  res
}

run_ancombc2 <- function(counts_df_samples_x_features, meta_df, group_col) {
  stopifnot(requireNamespace("ANCOMBC", quietly = TRUE))

  otu <- t(as.matrix(counts_df_samples_x_features)) # taxa x samples
  meta <- meta_df
  meta[[group_col]] <- as.factor(meta[[group_col]])

  fit <- ANCOMBC::ancombc2(
    data = otu,
    meta_data = meta,
    fix_formula = group_col,
    rand_formula = NULL,
    p_adj_method = "BH",
    prv_cut = 0, lib_cut = 0,
    group = group_col,
    struc_zero = TRUE,
    neg_lb = TRUE,
    alpha = 0.05,
    global = TRUE
  )

  out <- list(fit = fit, res = NULL, n_sig = NA_integer_)
  if (!is.null(fit$res) && is.data.frame(fit$res)) {
    out$res <- fit$res
    diff_cols <- grep("^diff_", colnames(out$res), value = TRUE)
    if (length(diff_cols) > 0) out$n_sig <- sum(out$res[, diff_cols] == TRUE, na.rm = TRUE)
  }
  out
}

run_da_suite <- function(counts_df_samples_x_features, meta_sub, group_col, output_dir, domain_name) {
  suppressPackageStartupMessages(library(dplyr))
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

  x <- counts_df_samples_x_features[rownames(meta_sub), , drop = FALSE]
  x <- x[, colSums(x) > 0, drop = FALSE]

  meta_sub[[group_col]] <- as.factor(meta_sub[[group_col]])
  ng <- nlevels(meta_sub[[group_col]])

  cat("\n=== Running DA Suite for:", domain_name, "===\n")
  cat("Groups:", paste(levels(meta_sub[[group_col]]), collapse = ", "),
    " (n=", nrow(meta_sub), ")\n",
    sep = ""
  )
  cat("Taxa/features tested:", ncol(x), "\n")

  # ALDEx2: ONLY for pairwise contrasts
  aldex_res <- NULL
  if (requireNamespace("ALDEx2", quietly = TRUE) && ng == 2) {
    counts_mat <- t(as.matrix(x)) # features x samples
    aldex_res <- tryCatch(
      run_aldex2_pairwise(counts_mat, meta_sub[[group_col]]),
      error = function(e) {
        message("ALDEx2 failed: ", conditionMessage(e))
        NULL
      }
    )
    if (!is.null(aldex_res)) {
      saveRDS(aldex_res, file.path(output_dir, "aldex2_results.rds"))
      write.csv(aldex_res, file.path(output_dir, "aldex2_results.csv"), row.names = FALSE)
    }
  }

  # ANCOM-BC2 always
  ancom_res <- NULL
  ancom_global_tbl <- NULL
  if (requireNamespace("ANCOMBC", quietly = TRUE)) {
    ancom_res <- tryCatch(
      run_ancombc2(x, meta_sub, group_col),
      error = function(e) {
        message("ANCOM-BC2 failed: ", conditionMessage(e))
        NULL
      }
    )

    if (!is.null(ancom_res) && !is.null(ancom_res$res) && is.data.frame(ancom_res$res)) {
      saveRDS(ancom_res$res, file.path(output_dir, "ancombc2_results.rds"))
      write.csv(ancom_res$res, file.path(output_dir, "ancombc2_results.csv"), row.names = FALSE)
    }

    if (!is.null(ancom_res) && !is.null(ancom_res$fit) && !is.null(ancom_res$fit$res_global)) {
      ancom_global_tbl <- ancom_res$fit$res_global
      if (is.data.frame(ancom_global_tbl) && !("taxon" %in% colnames(ancom_global_tbl))) {
        ancom_global_tbl$taxon <- rownames(ancom_global_tbl)
        rownames(ancom_global_tbl) <- NULL
      }
      saveRDS(ancom_global_tbl, file.path(output_dir, "ancombc2_global.rds"))
      write.csv(ancom_global_tbl, file.path(output_dir, "ancombc2_global.csv"), row.names = FALSE)
    }
  }

  list(aldex = aldex_res, ancom = ancom_res, ancom_global = ancom_global_tbl)
}
```

---

# 7) CLR heatmap + per-taxon boxplots (top 5–10)

```{r clr_heatmap_boxplots}
# Heatmap: uses ggplot tile (no extra dependencies)
plot_clr_heatmap <- function(counts_df, meta_sub, feature_ids, group_col, title,
                             scale_rows = TRUE, max_features = 40) {
  suppressPackageStartupMessages({
    library(ggplot2)
    library(dplyr)
    library(tidyr)
    library(tibble)
  })

  feats <- unique(feature_ids)
  feats <- feats[feats %in% colnames(counts_df)]
  feats <- head(feats, max_features)
  if (length(feats) < 2) {
    return(NULL)
  }

  clr <- clr_transform(counts_df[rownames(meta_sub), feats, drop = FALSE])
  mat <- as.matrix(clr)

  if (scale_rows) {
    mat <- t(scale(t(mat)))
    mat[is.na(mat)] <- 0
  }

  df <- as.data.frame(mat) %>%
    tibble::rownames_to_column("SampleID_row") %>%
    tidyr::pivot_longer(-SampleID_row, names_to = "Feature", values_to = "CLR")

  meta_join <- meta_sub %>%
    tibble::rownames_to_column("SampleID_row")

  df <- df %>%
    dplyr::left_join(meta_join, by = "SampleID_row")

  ggplot(df, aes(x = Feature, y = SampleID_row, fill = CLR)) +
    geom_tile() +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
      axis.text.y = element_text(size = 8)
    ) +
    labs(title = title, x = "Feature", y = "Sample")
}

plot_top_taxa_boxplots <- function(counts_df, meta_sub, feature_ids, group_col, title, top_n = 10) {
  suppressPackageStartupMessages({
    library(ggplot2)
    library(dplyr)
    library(tidyr)
    library(tibble)
  })

  feats <- unique(feature_ids)
  feats <- feats[feats %in% colnames(counts_df)]
  feats <- head(feats, top_n)
  if (length(feats) < 1) {
    return(NULL)
  }

  clr <- clr_transform(counts_df[rownames(meta_sub), feats, drop = FALSE])

  df <- as.data.frame(clr) %>%
    tibble::rownames_to_column("SampleID_row") %>%
    tidyr::pivot_longer(-SampleID_row, names_to = "Feature", values_to = "CLR")

  meta_join <- meta_sub %>%
    tibble::rownames_to_column("SampleID_row")

  df <- df %>%
    dplyr::left_join(meta_join, by = "SampleID_row")

  ggplot(df, aes(x = .data[[group_col]], y = CLR)) +
    geom_boxplot(outlier.size = 0.6) +
    geom_jitter(width = 0.15, alpha = 0.7, size = 2) +
    facet_wrap(~Feature, scales = "free_y") +
    theme_bw() +
    labs(title = title, x = group_col, y = "CLR abundance")
}
```

---

# 8) Domain runner (prints tables + inline plots + saves PDFs) + stratified PERMANOVA

```{r domain_runner}
run_domain_for_all_contrasts <- function(counts_df_samples_x_features, domain_name, metadata, contrasts, out_dir, nperm = 999) {
  suppressPackageStartupMessages({
    library(dplyr)
    library(knitr)
    library(ggplot2)
  })
  results <- list()

  for (cc in contrasts) {
    meta_sub <- metadata %>% filter(!!cc$filter_expr)
    meta_sub <- meta_sub[!is.na(meta_sub[[cc$design_col]]), , drop = FALSE]

    # Add Orchard column (heuristic)
    meta_sub$Orchard <- infer_orchard(rownames(meta_sub))

    common <- intersect(rownames(counts_df_samples_x_features), rownames(meta_sub))
    if (length(common) < 5) next
    meta_sub <- meta_sub[common, , drop = FALSE]

    qc_before <- summarize_matrix_qc(counts_df_samples_x_features, meta_sub)

    fp <- domain_filter_params(domain_name, nrow(meta_sub))
    counts_f <- filter_counts(
      counts_df_samples_x_features, meta_sub,
      min_prev_frac = fp$min_prev_frac,
      min_total = fp$min_total
    )

    qc_after <- data.frame(
      n_samples = nrow(meta_sub),
      n_features = ncol(counts_f),
      min_prev_frac_used = fp$min_prev_frac,
      min_total_used = fp$min_total
    )

    contrast_dir <- file.path(out_dir, domain_name, cc$name)
    dir.create(contrast_dir, recursive = TRUE, showWarnings = FALSE)
    write.csv(qc_before, file.path(contrast_dir, "qc_before_filtering.csv"), row.names = FALSE)
    write.csv(qc_after, file.path(contrast_dir, "qc_after_filtering.csv"), row.names = FALSE)

    da <- run_da_suite(
      counts_df_samples_x_features = counts_f,
      meta_sub = meta_sub,
      group_col = cc$design_col,
      output_dir = contrast_dir,
      domain_name = paste0(domain_name, "_", cc$name)
    )

    # Prevalence enrichment (pairwise only)
    prev_res <- NULL
    if (!is.null(cc$levels) && length(cc$levels) == 2) {
      prev_res <- run_prevalence_enrichment(
        counts_df = counts_f, meta_sub = meta_sub,
        group_col = cc$design_col, g1 = cc$levels[1], g2 = cc$levels[2],
        detect_thresh = 1
      )
      if (!is.null(prev_res)) {
        saveRDS(prev_res, file.path(contrast_dir, "prevalence_enrichment.rds"))
        write.csv(prev_res, file.path(contrast_dir, "prevalence_enrichment.csv"), row.names = FALSE)
      }
    }

    # PERMANOVA + dispersion
    perm <- run_aitchison_permanova(
      counts_df = counts_f, meta_sub = meta_sub, design_col = cc$design_col,
      title_tag = paste0(domain_name, "_", cc$name),
      output_dir = contrast_dir, nperm = nperm
    )

    # Optional: orchard-stratified PERMANOVA (reviewer request)
    perm_str <- run_stratified_permanova(
      counts_df = counts_f, meta_sub = meta_sub,
      design_col = cc$design_col, strata_col = "Orchard",
      title_tag = paste0(domain_name, "_", cc$name),
      output_dir = contrast_dir, nperm = nperm
    )

    # ---------- Inline reporting ----------
    cat("\n\n## ", domain_name, " — ", cc$name, "\n", sep = "")
    cat("Design: ", cc$design_col,
      " | n_samples=", qc_after$n_samples,
      " | n_features_tested=", qc_after$n_features, "\n",
      sep = ""
    )
    if (!is.null(cc$note)) cat("Note: ", cc$note, "\n", sep = "")

    if (cc$name == "Global_Group4") {
      cat("\nNOTE: Global_Group4 is descriptive only (composite groups; confounded).\n")
      cat("Feature-level DA should be treated as exploratory.\n")
    }

    cat("\nQC before filtering:\n")
    print(knitr::kable(qc_before, digits = 3))
    cat("\nQC after filtering:\n")
    print(knitr::kable(qc_after, digits = 3))

    # PCoA (inline + saved)
    p_pcoa <- plot_aitchison_pcoa(
      counts_df = counts_f, meta_sub = meta_sub,
      color_col = cc$design_col,
      title = paste0(pretty_domain(domain_name), " \u2014 ", pretty_contrast(cc$name), " (Aitchison PCoA)"),
      palette_named = if (cc$design_col == "Group4") group4_colors else NULL
    )
    if (!is.null(p_pcoa)) {
      print(p_pcoa)
      ggsave(file.path(contrast_dir, "pcoa_aitchison.pdf"), p_pcoa, width = 8, height = 6)
    }

    # PERMANOVA outputs
    if (!is.null(perm)) {
      cat("\n### PERMANOVA (adonis2)\n")
      print(perm$adonis)
      cat("\n### Dispersion (betadisper permutest)\n")
      print(perm$betadisper)
      cat("\n(perm requested=", perm$nperm_requested, "; vegan used the number printed above)\n", sep = "")
    }

    if (!is.null(perm_str)) {
      cat("\n### PERMANOVA stratified by Orchard (reviewer-style robustness)\n")
      cat("Orchard strata levels:", length(unique(meta_sub$Orchard)), "\n")
      print(perm_str$adonis_stratified)
    } else {
      cat("\n### PERMANOVA stratified by Orchard: not run (insufficient strata/replication in this subset).\n")
    }

    # ALDEx2 (pairwise only)
    if (!is.null(da$aldex) && is.data.frame(da$aldex) && nrow(da$aldex) > 0) {
      cat("\n### ALDEx2 (pairwise): top 15 by wi.eBH\n")
      print(knitr::kable(head(da$aldex[order(da$aldex$wi.eBH), ], 15), digits = 3))
    } else {
      cat("\n### ALDEx2: not run (only valid for 2-level contrasts) or no results.\n")
    }

    # ANCOM-BC2 (non-intercept coefficient)
    sig_features_for_plots <- character(0)

    if (!is.null(da$ancom) && !is.null(da$ancom$res) && is.data.frame(da$ancom$res)) {
      cat("\n### ANCOM-BC2 available coefficient columns\n")
      print(colnames(da$ancom$res))

      coef <- pick_ancom_coef(da$ancom$res)
      q_col <- paste0("q_", coef)
      lfc_col <- paste0("lfc_", coef)
      diff_col <- paste0("diff_", coef)
      diff_rob_col <- paste0("diff_robust_", coef)

      if (q_col %in% colnames(da$ancom$res) && lfc_col %in% colnames(da$ancom$res)) {
        cat("\n### ANCOM-BC2: top 15 by ", q_col, " (non-intercept)\n", sep = "")
        top_an <- da$ancom$res[order(da$ancom$res[[q_col]]), ]
        print(knitr::kable(head(top_an, 15), digits = 3))

        sig_an <- da$ancom$res[
          !is.na(da$ancom$res[[q_col]]) &
            da$ancom$res[[q_col]] < 0.05 &
            ((da$ancom$res[[diff_col]] == TRUE) | (da$ancom$res[[diff_rob_col]] == TRUE)), ,
          drop = FALSE
        ]
        sig_an <- sig_an[order(sig_an[[q_col]]), , drop = FALSE]

        if (nrow(sig_an) > 0) {
          cat("\n### ANCOM-BC2 significant: q<0.05 for ", coef, "\n", sep = "")
          print(knitr::kable(head(sig_an, 25), digits = 3))
          sig_features_for_plots <- sig_an$taxon
        } else {
          cat("\n### ANCOM-BC2 significant: none at q<0.05 for ", coef, "\n", sep = "")
        }

        p_ancom <- plot_ancombc_bars(da$ancom, paste0(pretty_domain(domain_name), " \u2014 ", pretty_contrast(cc$name)), contrast_level = coef, top_n = 20)
        if (!is.null(p_ancom)) {
          print(p_ancom)
          ggsave(file.path(contrast_dir, paste0("ancombc2_lfc_", coef, ".pdf")), p_ancom, width = 10, height = 7)
        }
      }
    }

    # ANCOM global omnibus (useful esp. for multi-group)
    if (!is.null(da$ancom_global) && is.data.frame(da$ancom_global)) {
      cat("\n### ANCOM-BC2 Global test (top 15 by q_val)\n")
      gtab <- da$ancom_global[order(da$ancom_global$q_val), ]
      print(knitr::kable(head(gtab, 15), digits = 3))
    }

    # Prevalence enrichment + plot
    if (!is.null(prev_res) && is.data.frame(prev_res) && nrow(prev_res) > 0) {
      cat("\n### Prevalence enrichment (top 15 by q)\n")
      print(knitr::kable(head(prev_res, 15), digits = 3))

      p_prev <- plot_prevalence_shift(prev_res, paste0(pretty_domain(domain_name), " \u2014 ", pretty_contrast(cc$name), " (Prevalence Shift)"), top_n = 20)
      if (!is.null(p_prev)) {
        print(p_prev)
        ggsave(file.path(contrast_dir, "prevalence_shift.pdf"), p_prev, width = 9, height = 7)
      }
    }

    # CLR heatmap + boxplots for significant ANCOM taxa (if any)
    if (length(sig_features_for_plots) > 0) {
      # Heatmap (top 40 sig)
      p_hm <- plot_clr_heatmap(
        counts_df = counts_f, meta_sub = meta_sub,
        feature_ids = sig_features_for_plots,
        group_col = cc$design_col,
        title = paste0(pretty_domain(domain_name), " \u2014 ", pretty_contrast(cc$name), " (CLR Abundance Heatmap)"),
        scale_rows = TRUE, max_features = 40
      )
      if (!is.null(p_hm)) {
        print(p_hm)
        ggsave(file.path(contrast_dir, "clr_heatmap_sig_ancom.pdf"), p_hm, width = 12, height = 7)
      }

      # Boxplots (top 10 sig)
      p_box <- plot_top_taxa_boxplots(
        counts_df = counts_f, meta_sub = meta_sub,
        feature_ids = sig_features_for_plots,
        group_col = cc$design_col,
        title = paste0(pretty_domain(domain_name), " \u2014 ", pretty_contrast(cc$name), " (Top Differentially Abundant Taxa)"),
        top_n = 10
      )
      if (!is.null(p_box)) {
        print(p_box)
        ggsave(file.path(contrast_dir, "clr_boxplots_top_sig_ancom.pdf"), p_box, width = 12, height = 7)
      }
    } else {
      cat("\n### CLR heatmap/boxplots: skipped (no significant ANCOM taxa at q<0.05).\n")
    }

    results[[cc$name]] <- list(
      contrast = cc,
      meta = meta_sub,
      counts_filtered = counts_f,
      da = da,
      prevalence = prev_res,
      permanova = perm,
      permanova_stratified = perm_str,
      qc_before = qc_before,
      qc_after = qc_after,
      contrast_dir = contrast_dir
    )
  }

  results
}
```

---

# 9) Run all domains

```{r run_all_domains}
NPERM <- 999

prok_results <- run_domain_for_all_contrasts(
  counts_df_samples_x_features = mag_counts,
  domain_name = "Prokaryotes_MAG_species",
  metadata = metadata,
  contrasts = contrasts,
  out_dir = OUT_DIR,
  nperm = NPERM
)

plasmid_results <- if (!is.null(plasmid_counts)) {
  run_domain_for_all_contrasts(
    counts_df_samples_x_features = plasmid_counts,
    domain_name = "Plasmids_host_family",
    metadata = metadata,
    contrasts = contrasts,
    out_dir = OUT_DIR,
    nperm = NPERM
  )
} else {
  NULL
}

virus_results <- if (!is.null(virus_counts)) {
  run_domain_for_all_contrasts(
    counts_df_samples_x_features = virus_counts,
    domain_name = "Viruses_vOTU",
    metadata = metadata,
    contrasts = contrasts,
    out_dir = OUT_DIR,
    nperm = NPERM
  )
} else {
  NULL
}
```

---

# 10) Summary table

```{r da_summary}
suppressPackageStartupMessages(library(dplyr))

get_aldex_sig <- function(res) {
  if (is.null(res) || !is.data.frame(res) || nrow(res) == 0) {
    return(NA_integer_)
  }
  if ("wi.eBH" %in% colnames(res)) {
    return(sum(res$wi.eBH < 0.05, na.rm = TRUE))
  }
  NA_integer_
}

get_ancom_sig <- function(obj) {
  if (is.null(obj) || is.null(obj$res) || !is.data.frame(obj$res)) {
    return(NA_integer_)
  }
  diff_cols <- grep("^diff_", colnames(obj$res), value = TRUE)
  if (length(diff_cols) == 0) {
    return(NA_integer_)
  }
  sum(obj$res[, diff_cols] == TRUE, na.rm = TRUE)
}

summarize_domain <- function(domain_res, domain_label) {
  if (is.null(domain_res)) {
    return(NULL)
  }

  rows <- lapply(names(domain_res), function(nm) {
    obj <- domain_res[[nm]]
    da <- obj$da

    coef_used <- NA_character_
    ancom_sig_coef <- NA_integer_

    if (!is.null(da$ancom) && !is.null(da$ancom$res) && is.data.frame(da$ancom$res)) {
      coef_used <- pick_ancom_coef(da$ancom$res)
      q_col <- paste0("q_", coef_used)
      diff_col <- paste0("diff_", coef_used)
      diff_rob_col <- paste0("diff_robust_", coef_used)
      if (q_col %in% colnames(da$ancom$res)) {
        ancom_sig_coef <- sum(
          da$ancom$res[[q_col]] < 0.05 &
            ((da$ancom$res[[diff_col]] == TRUE) | (da$ancom$res[[diff_rob_col]] == TRUE)),
          na.rm = TRUE
        )
      }
    }

    data.frame(
      Domain = domain_label,
      Contrast = nm,
      Design = obj$contrast$design_col,
      n_samples = obj$qc_after$n_samples,
      n_features_tested = obj$qc_after$n_features,
      ALDEx2_sig = get_aldex_sig(da$aldex),
      ANCOMBC2_sig_anycoef = get_ancom_sig(da$ancom),
      ANCOMBC2_coef_used = coef_used,
      ANCOMBC2_sig_coef = ancom_sig_coef,
      ANCOMBC2_global_sig = if (!is.null(da$ancom_global) && is.data.frame(da$ancom_global)) {
        sum(da$ancom_global$diff_abn == TRUE, na.rm = TRUE)
      } else {
        NA_integer_
      },
      PrevEnrich_sig = if (!is.null(obj$prevalence) && is.data.frame(obj$prevalence)) {
        sum(obj$prevalence$q < 0.05, na.rm = TRUE)
      } else {
        NA_integer_
      },
      Stratified_PERMANOVA_run = !is.null(obj$permanova_stratified)
    )
  })

  bind_rows(rows)
}

summary_all <- bind_rows(
  summarize_domain(prok_results, "Prokaryotes_MAG_species"),
  summarize_domain(plasmid_results, "Plasmids_host_family"),
  summarize_domain(virus_results, "Viruses_vOTU")
)

knitr::kable(summary_all, caption = "DA Summary (ANCOM-BC2 + ALDEx2 pairwise + prevalence + stratified PERMANOVA)", digits = 3)
write.csv(summary_all, file.path(OUT_DIR, "da_summary_all_domains.csv"), row.names = FALSE)
```

---

# Session info

```{r session}
sessionInfo()
```