---
title: "07 - Community–Resistome Coupling Analysis"
author: "Analysis Pipeline"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE, warning = FALSE, message = FALSE,
    fig.width = 10, fig.height = 7
)
```

# Overview

This notebook tests whether **prokaryotic community composition** (MAGs) is
coupled to the **contig-level resistome** (ARG drug-class and mechanism
profiles). Coupling is assessed via:
1. **Mantel test** — correlation between Aitchison distance matrices
2. **Procrustes analysis** — superimposition of PCoA ordinations

All distances are **Aitchison** (CLR + Euclidean), consistent with the
compositional framework used throughout the project.

---

# 1. Setup and Data Loading

```{r load_setup}
source("/media/Box/applesurface_management/00_Data_antigravity/scripts/00_setup.R")

OUT_DIR <- file.path(RESULTS_DIR, "07_community_resistome_coupling")
if (!dir.exists(OUT_DIR)) dir.create(OUT_DIR, recursive = TRUE)

cat("Output directory:", OUT_DIR, "\n")
```

## Load Community and Resistome Matrices

```{r load_data}
# ---- Community Matrices ----
mag_genus <- readRDS(file.path(RESULTS_DIR, "data_objects", "mag_genus_counts_filt.rds"))
mag_species <- readRDS(file.path(RESULTS_DIR, "data_objects", "mag_species_counts_filt.rds"))

# ---- Resistome Matrices ----
arg_class <- readRDS(file.path(RESULTS_DIR, "data_objects", "arg_mat_contig.rds"))
arg_mech <- readRDS(file.path(RESULTS_DIR, "data_objects", "arg_mech_mat_contig.rds"))

cat("MAG genus:     ", nrow(mag_genus), "samples ×", ncol(mag_genus), "taxa\n")
cat("MAG species:   ", nrow(mag_species), "samples ×", ncol(mag_species), "taxa\n")
cat("ARG drug class:", nrow(arg_class), "samples ×", ncol(arg_class), "classes\n")
cat("ARG mechanism: ", nrow(arg_mech), "samples ×", ncol(arg_mech), "mechanisms\n")
```

---

# 2. Define Comparisons

```{r define_comparisons}
# Define the community–resistome pairings to test
comparisons <- list(
    list(
        name        = "MAG Genus vs ARG Drug Class",
        community   = mag_genus,
        resistome   = arg_class,
        comm_label  = "MAG Genus",
        res_label   = "ARG Drug Class",
        short_name  = "mag_genus_vs_arg_class"
    ),
    list(
        name        = "MAG Genus vs ARG Mechanism",
        community   = mag_genus,
        resistome   = arg_mech,
        comm_label  = "MAG Genus",
        res_label   = "ARG Mechanism",
        short_name  = "mag_genus_vs_arg_mech"
    ),
    list(
        name        = "MAG Species vs ARG Drug Class",
        community   = mag_species,
        resistome   = arg_class,
        comm_label  = "MAG Species",
        res_label   = "ARG Drug Class",
        short_name  = "mag_species_vs_arg_class"
    ),
    list(
        name        = "MAG Species vs ARG Mechanism",
        community   = mag_species,
        resistome   = arg_mech,
        comm_label  = "MAG Species",
        res_label   = "ARG Mechanism",
        short_name  = "mag_species_vs_arg_mech"
    )
)

cat("Defined", length(comparisons), "community–resistome pairings:\n")
for (cc in comparisons) cat(" •", cc$name, "\n")
```

---

# 3. Coupling Analysis

```{r run_coupling, fig.height=8, fig.width=10, results='asis'}
# ---- Helper: run full coupling pipeline for one comparison ----
run_coupling <- function(comp, metadata, group4_colors, out_dir) {
    library(vegan)
    library(ggrepel)

    # 1. Match samples

    common <- intersect(rownames(comp$community), rownames(comp$resistome))
    if (length(common) < 3) {
        cat("\n**", comp$name, "** — Skipped (only", length(common), "shared samples)\n\n")
        return(NULL)
    }

    comm_sub <- comp$community[common, , drop = FALSE]
    res_sub <- comp$resistome[common, , drop = FALSE]

    # 2. Filter low-prevalence features
    comm_filt <- filter_features(comm_sub)
    res_filt <- filter_features(res_sub)

    if (ncol(comm_filt) < 2 || ncol(res_filt) < 2) {
        cat("\n**", comp$name, "** — Skipped (too few features after filtering)\n\n")
        return(NULL)
    }

    # 3. CLR transform
    comm_clr <- calc_clr(comm_filt)
    res_clr <- calc_clr(res_filt)

    # 4. Aitchison distance
    comm_dist <- calc_aitchison(comm_clr)
    res_dist <- calc_aitchison(res_clr)

    # ============================
    # 5. MANTEL TEST (Spearman)
    # ============================
    mantel_res <- vegan::mantel(comm_dist, res_dist,
        method = "spearman", permutations = 9999
    )

    # ============================
    # 6. PROCRUSTES + PROTEST
    # ============================
    comm_pcoa <- cmdscale(comm_dist, k = 2) # PCoA (classical MDS)
    res_pcoa <- cmdscale(res_dist, k = 2)

    proc <- vegan::procrustes(comm_pcoa, res_pcoa)
    proc_test <- vegan::protest(comm_pcoa, res_pcoa, permutations = 9999)

    # ============================
    # 7. Build results list
    # ============================
    result <- list(
        name              = comp$name,
        short_name        = comp$short_name,
        n_samples         = length(common),
        n_comm_features   = ncol(comm_filt),
        n_res_features    = ncol(res_filt),
        mantel_r          = mantel_res$statistic,
        mantel_p          = mantel_res$signif,
        mantel_method     = "Spearman",
        mantel_perms      = 9999,
        procrustes_m2     = proc_test$ss,
        procrustes_corr   = proc_test$t0,
        procrustes_p      = proc_test$signif,
        procrustes_perms  = 9999,
        comm_label        = comp$comm_label,
        res_label         = comp$res_label,
        proc              = proc,
        proc_test         = proc_test,
        comm_pcoa         = comm_pcoa,
        res_pcoa          = res_pcoa,
        comm_dist         = comm_dist,
        res_dist          = res_dist,
        common_samples    = common
    )

    # ============================
    # 8. Pretty Procrustes Plot
    # ============================
    # Build data frame for ggplot
    target_df <- as.data.frame(proc$Yrot) # Target (rotated community)
    rotated_df <- as.data.frame(proc$X) # Rotated resistome

    colnames(target_df) <- c("tPC1", "tPC2")
    colnames(rotated_df) <- c("rPC1", "rPC2")

    target_df$Sample <- rownames(target_df)
    rotated_df$Sample <- rownames(rotated_df)

    plot_df <- merge(target_df, rotated_df, by = "Sample")
    plot_df <- merge(plot_df, metadata, by.x = "Sample", by.y = "SampleID")

    # Annotation text
    annot_text <- paste0(
        "Mantel r = ", round(result$mantel_r, 3),
        " (p = ", format.pval(result$mantel_p, digits = 3), ")\n",
        "Procrustes m² = ", round(result$procrustes_m2, 3),
        " (p = ", format.pval(result$procrustes_p, digits = 3), ")"
    )

    p_proc <- ggplot(plot_df) +
        # Arrows from community to resistome
        geom_segment(aes(x = rPC1, y = rPC2, xend = tPC1, yend = tPC2),
            arrow = arrow(length = unit(0.15, "cm"), type = "closed"),
            color = "grey50", linewidth = 0.4, alpha = 0.7
        ) +
        # Community points (circles)
        geom_point(aes(x = rPC1, y = rPC2, fill = Group4),
            shape = 21, size = 4, stroke = 0.5
        ) +
        # Resistome points (triangles)
        geom_point(aes(x = tPC1, y = tPC2, fill = Group4),
            shape = 24, size = 4, stroke = 0.5
        ) +
        # Labels
        geom_text_repel(aes(x = rPC1, y = rPC2, label = Sample),
            size = 2.5, max.overlaps = 20, color = "grey30"
        ) +
        scale_fill_manual(values = group4_colors) +
        # Statistics annotation
        annotate("text",
            x = Inf, y = -Inf, label = annot_text,
            hjust = 1.05, vjust = -0.3, size = 3.5, fontface = "italic"
        ) +
        theme_bw(base_size = 13) +
        labs(
            title = paste("Procrustes:", comp$comm_label, "↔", comp$res_label),
            subtitle = paste0(
                "n = ", result$n_samples, " samples | ",
                result$n_comm_features, " community features, ",
                result$n_res_features, " resistome features"
            ),
            x = "Dimension 1",
            y = "Dimension 2",
            fill = "Management Group",
            caption = paste0("●  = ", comp$comm_label, " (community)    ▲  = ", comp$res_label, " (resistome)")
        ) +
        theme(
            plot.caption = element_text(hjust = 0.5, size = 10),
            legend.position = "right"
        )

    # Save
    ggsave(file.path(out_dir, paste0("procrustes_", comp$short_name, ".pdf")),
        p_proc,
        width = 10, height = 7
    )
    print(p_proc)

    # ============================
    # 9. Distance Matrix Scatter
    # ============================
    # Extract lower-triangle distances
    comm_vec <- as.vector(comm_dist)
    res_vec <- as.vector(res_dist)

    dist_df <- data.frame(Community = comm_vec, Resistome = res_vec)

    p_scatter <- ggplot(dist_df, aes(x = Community, y = Resistome)) +
        geom_point(alpha = 0.6, size = 2, color = "#2166AC") +
        geom_smooth(method = "lm", se = TRUE, color = "#B2182B", linewidth = 0.8) +
        annotate("text",
            x = Inf, y = -Inf, hjust = 1.05, vjust = -0.5,
            label = paste0(
                "Mantel r = ", round(result$mantel_r, 3),
                "\np = ", format.pval(result$mantel_p, digits = 3)
            ),
            size = 4, fontface = "italic"
        ) +
        theme_bw(base_size = 13) +
        labs(
            title = paste("Distance Matrix Correlation:", comp$comm_label, "vs", comp$res_label),
            x = paste(comp$comm_label, "— Aitchison Distance"),
            y = paste(comp$res_label, "— Aitchison Distance")
        )

    ggsave(file.path(out_dir, paste0("mantel_scatter_", comp$short_name, ".pdf")),
        p_scatter,
        width = 8, height = 6
    )
    print(p_scatter)

    return(result)
}

# ---- Run all comparisons ----
all_results <- list()

for (cc in comparisons) {
    cat("\n\n## ", cc$name, "\n\n", sep = "")
    res <- run_coupling(cc, metadata, group4_colors, OUT_DIR)
    if (!is.null(res)) {
        all_results[[cc$short_name]] <- res
    }
}
```

---

# 4. Summary Table

```{r summary_table}
if (length(all_results) > 0) {
    summary_df <- data.frame(
        Comparison         = sapply(all_results, `[[`, "name"),
        Samples            = sapply(all_results, `[[`, "n_samples"),
        Community_Features = sapply(all_results, `[[`, "n_comm_features"),
        Resistome_Features = sapply(all_results, `[[`, "n_res_features"),
        Distance_Metric    = "Aitchison (CLR + Euclidean)",
        Mantel_r           = round(sapply(all_results, `[[`, "mantel_r"), 4),
        Mantel_p           = sapply(all_results, function(x) format.pval(x$mantel_p, digits = 3)),
        Mantel_Method      = "Spearman",
        Mantel_Perms       = sapply(all_results, `[[`, "mantel_perms"),
        Procrustes_m2      = round(sapply(all_results, `[[`, "procrustes_m2"), 4),
        Procrustes_corr    = round(sapply(all_results, `[[`, "procrustes_corr"), 4),
        Procrustes_p       = sapply(all_results, function(x) format.pval(x$procrustes_p, digits = 3)),
        Procrustes_Perms   = sapply(all_results, `[[`, "procrustes_perms"),
        stringsAsFactors   = FALSE
    )
    rownames(summary_df) <- NULL

    # Print
    cat("\n=== Community–Resistome Coupling Summary ===\n\n")
    print(knitr::kable(summary_df, format = "pipe", align = "l"))

    # Save
    write.csv(summary_df, file.path(OUT_DIR, "coupling_summary.csv"), row.names = FALSE)
    cat("\nSaved to:", file.path(OUT_DIR, "coupling_summary.csv"), "\n")
}
```

---

# 5. Interpretation Guide

```{r interpretation, results='asis'}
cat("
### How to Read These Results

| Metric | Interpretation |
|:-------|:---------------|
| **Mantel r** | Correlation between distance matrices. Range: −1 to 1. Values > 0.3 suggest moderate coupling. |
| **Mantel p** | Permutation p-value. Significant if < 0.05. |
| **Procrustes m²** | Sum of squared residuals. **Lower = stronger concordance.** Range: 0 to 1. |
| **Procrustes correlation** | Correlation in symmetric Procrustes rotation. **Higher = better fit.** |
| **Procrustes p** | Permutation p-value from `protest()`. Significant if < 0.05. |

**Distance metric:** Aitchison (CLR transformation with multiplicative zero replacement → Euclidean distance), chosen for its compositionality-aware properties.

**Correlation method:** Spearman rank correlation (non-parametric, robust to outliers).

**Permutations:** 9,999 per test, providing resolution down to p = 0.0001.

### What the Procrustes plots show

- **Circles (●)** = community ordination (MAG genus or species)
- **Triangles (▲)** = resistome ordination (ARG drug class or mechanism)
- **Arrows** connect the **same sample** in both ordinations
- **Shorter arrows** = stronger agreement between community and resistome structure for that sample
- Colors distinguish Management × Country groups
")
```

---

# 6. Save Results

```{r save_results}
if (length(all_results) > 0) {
    # Save full results (without bulky objects)
    save_results <- lapply(all_results, function(x) {
        x$proc <- NULL
        x$proc_test <- NULL
        x$comm_pcoa <- NULL
        x$res_pcoa <- NULL
        x$comm_dist <- NULL
        x$res_dist <- NULL
        x
    })
    saveRDS(save_results, file.path(OUT_DIR, "coupling_results.rds"))
    cat("Results saved to:", OUT_DIR, "\n")
}
```

---

# Session Info

```{r session}
sessionInfo()
```
